---
title: "Search Species in PEZ"
author: Synthesis prepared for the CSAS Aquaculture Siting Review by the MSP Science
  Team
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r Define Aquaculture Site, echo=FALSE, include=FALSE, cache=FALSE}
rm(list=ls(all=TRUE))
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"

#make sure you delete all shp.xml files in the folder to be able to run this code
minYear <- 2002
cat(AquaSiteName)
```
```{r setup, echo=FALSE, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(Mar.datawrangling)
library(knitr)
library(kableExtra)
library(rgdal)
library(maps)
library(lubridate)
library(RCurl)
#source("C:/RProjects/SearchPEZ/code/.Rprofile") #Catalina's path to access Oracle passwords
```
```{r source functions, echo=FALSE, include=FALSE, cache=FALSE}
# Here code reads from HES_MSP. However, code is in Catalina's computer and pushed to  https://github.com/AtlanticR/MSP/tree/master/SearchPEZ/code 
source("//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/code/site_map.r")
source("//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/code/plot_wsdb.r")
source("//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/code/filter_wsdb.r")
#source("//dcnsbiona01a/BIODataSvc/IN/MSP/Projects/Aquaculture/SearchPEZ/code/remove_landPoints.r")
```

```{r Set up paths, echo=FALSE, include=FALSE, cache=FALSE}
mspPath <- "//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ"
data.dir <- "//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/inputs/mar.wrangling"
dataPath <- file.path("//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/outputs",paste0(AquaSiteName,PEZversion))
setwd(dataPath)
polyPath <- "//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/inputs/polygons"
wsdbFile <- file.path("//ent.dfo-mpo.ca/ATLShares/Science/CESD/HES_MSP/R/SearchPEZ/inputs","WSDB_GLazin_AllSpecies_2018.csv")
```
```{r read baseline data for maps, echo=FALSE, include=FALSE, cache=FALSE}
landGDB <- "//dcnsbiona01a/BIODataSvc/IN/MSP/Projects/BaseGISData/Boundaries/Coast50K/Coastline50k.gdb"
land <- readOGR(dsn=landGDB,layer="Land50k_Maritimes",stringsAsFactors=F)
```
```{r Define PEZ!, echo=FALSE, include=FALSE, cache=FALSE}
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
```
###**Potential Exposure Zones (PEZs) polygon**

**Site:**
```{r echo=FALSE, results='asis'}
cat(AquaSiteName)
```

The aquaculture Science group led by Fred Page works with the Aquaculture site information, including current meter data, to calculate the Potential Exposure Zones (PEZ) of the aquaculture inputs (depositional feed/feces, advective/dispersive fish health treatments, etc.). This PEZ polygon is then used in this document to search regional databases (**2002 - 2018**) and summarize species that have been recorded. This information is used to determine if other, more site-specific, information is available on species presence for PEZ as a complement to the information provided by the proponent. Species lists generated in this synthesis is subsequently circulated to units in Maritimes Science to verify and compliment the information provided (e.g. spatial and temporal presence/absence of species in the area). 

Map below shoes the PEZ polygon (blue) used to search regional databases and summarize recorded species.


```{r Fig 1 PEZ polygon, fig.height=8, fig.width=11}
site_map(PEZ_poly,site,land,40)
```

# **Regional Databases**

```{r PEZ search, echo=FALSE, include=FALSE, cache=FALSE}
#This chunk of code should be run to do a re-extraction of data or to extract data using a new polygon.
#Note taht you need 64-bit ODBC driver to be able to extract MARFIS data. Mike provided data to Catalina on Sept 2019 and waiting to get Oracle 64 bit to do extraction myself.

# get_data(db = 'rv', data.dir = data.dir)
# GSINF = clip_by_poly(df = GSINF
#                       , lat.field = "LATITUDE"
#                       , lon.field = "LONGITUDE"
#                       , clip.poly = PEZ_poly)
# self_filter()
# save_data(db='rv')
# cleanup('rv')
# 
# get_data('isdb', data.dir=data.dir)
# ISSETPROFILE_WIDE = clip_by_poly(df = ISSETPROFILE_WIDE
#                              , lat.field = "LATITUDE"
#                              , lon.field = "LONGITUDE"
#                              , clip.poly = PEZ_poly)
# self_filter()
# save_data(db='isdb')
# cleanup('isdb')
# 
# get_data('marfis', data.dir = data.dir)
# PRO_SPC_INFO = clip_by_poly(df = PRO_SPC_INFO
#                       , lat.field = "LATITUDE"
#                       , lon.field = "LONGITUDE"
#                       , clip.poly = PEZ_poly)
# self_filter()
# save_data(db='marfis')
# cleanup('marfis')
```
```{r read csv extracted with previous chunck, echo=FALSE, include=FALSE, cache=FALSE}
fl <- list.files(dataPath,".csv")
marfis <- read.csv(file.path(dataPath,fl[grep("marfis_",fl)]))
marfis <- marfis[which(year(marfis$DATE_FISHED)>=minYear),]
isdb <- read.csv(file.path(dataPath,fl[grep("isdb_",fl)]))
isdb <- isdb[which(year(isdb$DATE_TIME2)>=minYear),]
rv <- read.csv(file.path(dataPath,fl[grep("rv_",fl)]))
rv <- rv[which(year(rv$SDATE)>=minYear),] 
```

### **Ecosystem Research Vessel (RV) Survey**

Fisheries and Oceans Canada (DFO) has conducted Research Vessel (RV) surveys in
the Maritimes Region using a standardized protocol. For more information see <https://waves-vagues.dfo-mpo.gc.ca/Library/4062092x.pdf>. The search of the Ecosystem RV Survey database resulted in the Table below that summarizes records by species or species group. 

```{r, echo=FALSE, include=FALSE, cache=FALSE}
rv$database <- "rv"
rv_freq <- aggregate(
        x = list(Records = rv$COMM),
      by = list(Species = rv$COMM
      ),
      length
      )
rv_freq <- rv_freq[order(-rv_freq$Records),]
rownames(rv_freq) <- c()
rv_total <- aggregate(
        x = list(total = rv$COMM),
      by = list(db = rv$database
      ),
      length
      )
Total_number_records_RV <- rv_total[[2]] 
Total_number_records_RV
```
```{r} 
kable(as.data.frame(Total_number_records_RV), digits=2, full_width = F, position = "left")
``` 
```{r}
kable(rv_freq) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
Map below shows the location of sets performed in the Ecosystem RV Survey database (records for all sets displayed are summarized in the table above). The yellow polygon shows the locations of the proposed aquaculture site.The blue polygon displays the PEZ used for this synthesis. 

```{r Fig 2 Ecosystem RV Survey records, fig.height=8, fig.width=11}
site_map(PEZ_poly,site,land,2)
points(rv$LONGITUDE,rv$LATITUDE, col="black",pch=16,cex=1.5)
```

### **Industry Survey Database (ISDB)**

ISDB is a DFO database that contains at-sea fish catch observations from commercial fishing vessels operating in the North West Atlantic.

```{r, echo=FALSE, include=FALSE, cache=FALSE}
isdb$database <- "isdb"
isdb_freq <- aggregate(
        x = list(Records = isdb$COMMON),
      by = list(Species = isdb$COMMON
      ),
      length
      )
isdb_freq <- isdb_freq[order(-isdb_freq$Records),]
rownames(isdb_freq) <- c()

isdb_total <- aggregate(
  x = list(total = isdb$COMMON),
  by = list(db = isdb$database
  ),
  length
)

Total_number_records_ISDB <- isdb_total[[2]] 
Total_number_records_ISDB
```

The search of ISDB resulted in the Table below that summarizes records by species or species group. Records were cropped to match PEZ.


```{r} 
kable(as.data.frame(Total_number_records_ISDB), digits=2, full_width = F, position = "left")
``` 

```{r}
kable(isdb_freq) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

### **The Maritime Fishery Information System (MARFIS)**

MARFIS is a DFO database, managed by the Policy and Economics Branch, that houses information on the fisheries of the Maritimes Region. This fishery monitoring information represents a complete census of almost all commercial fishing activities.  


```{r, echo=FALSE, include=FALSE, cache=FALSE}
marfis$database <- "marfis"
marfis_freq <- aggregate(
        x = list(Records = marfis$SPECIES_NAME),
      by = list(Species = marfis$SPECIES_NAME
      ),
      length
      )
marfis_freq <- marfis_freq[order(-marfis_freq$Records),]
rownames(marfis_freq) <- c()

marfis_total <- aggregate(
  x = list(total = marfis$COMM),
  by = list(db = marfis$database
  ),
  length
)
Total_number_records_MARFIS <- marfis_total[[2]] 
Total_number_records_MARFIS 
```
The search of MARFIS resulted in the Table below that summarizes records by species or species group. Records were cropped to match PEZ.


```{r}
kable(as.data.frame(Total_number_records_MARFIS), digits=2, full_width = F, position = "float_left") 
```
```{r}
kable(marfis_freq) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
Map below shows the location of samples recorded in the MARFIS and ISDB databases (records are summarized in the tables/sections above). The yellow polygon shows the locations of the proposed aquaculture site. The blue polygon displays the PEZ used for this synthesis. Records were cropped to match PEZ.

```{r Fig 3 MARFIS and ISDB records, fig.height=8, fig.width=11}
site_map(PEZ_poly,site,land,2)
points(marfis$ENT_LONGITUDE,marfis$ENT_LATITUDE, col="black",pch=16,cex=0.5)
points(isdb$LONG2,isdb$LAT2, col="black",pch=16,cex=0.5)
```

### **Whale Sightings Database **

Most sightings are collected on an opportunistic basis and observations may come from individuals with a variety of expertise in marine mammal identification experiences. Most data have been gathered from platforms of opportunity that were vessel-based. For more information please contact Pamela.Emery@dfo-mpo.gc.ca

Map below shows location of marine mammals sightings recorded in the Fisheries and Oceans whale sightings database. Map includes SARA and COSEWIC species within the region of interest. The yellow polygon shows the locations of the proposed aquaculture site. The blue polygon displays the PEZ used for this synthesis. 

```{r read wsdb and filter SARA & COSEWIC cetacean spp, echo=FALSE, include=FALSE, cache=FALSE}
wsdb <- read.csv(wsdbFile)
wsdb <- wsdb[which(year(as.Date(wsdb$WS_DATE,"%d-%b-%Y"))>=minYear),]
#wsdb <- remove_landPoints(wsdb,land)
wsdb_filter <- filter_wsdb(wsdb)
```
```{r Fig 4 Whale Sightings Database, fig.height=8, fig.width=11}
par(mfrow = c(1,2), mar=c(0, 0, 5.5, 0) + 2)#it goes c(bottom, left, top, right)
site_map(PEZ_poly,site,land,5)
plot_wsdb(PEZ_poly,wsdb_filter,5)
```

# **Resources**

### **Sources of Uncertainty**

Coastal areas are generally not adequately sampled on spatial and temporal scales of most relevance to aquaculture, i.e. tens to hundreds of meters and hours to months, and hence information on these space and time scales is generally not contained within the various data sources available to DFO, including the surveys referred to in this document. Therefore, there is uncertainty as to the exact distribution of species in the area of the proposed expansion.

MARFIS and ISDB do not fully sample the region spatially or temporally and, therefore, additional information on presence and habitat use (i.e. spawning, migration, feeding) must be drawn from larger-scale studies, which were also generally utilized by the proponent. 

The marine mammal sightings database has inherent problems with negative or positive reactions by cetaceans to the approach of such vessels that have not yet been factored into the data. Sighting effort has not been quantified (i.e., the numbers cannot be used to estimate true species density or abundance for an area). Lack of sightings do not represent lack of species present in a particular area. Numbers sighted have not been verified (especially in light of the significant differences in detectability among species). For completeness, the data represent an amalgamation of sightings from a variety of years and seasons. 

### **This document**

This synthesis is created using Rmarkdown and therefore this document can be reproduced and reused using the code  available in <https://github.com/AtlanticR/SearchPEZ>. The full document can be updated rapidly as more or different information becomes available.

This document was prepared by the Science MSP Team: <Catalina.Gomez@dfo-mpo.gc.ca>, <Gordana.Lazin@dfo-mpo.gc.ca>, <Brian.Bower@dfo-mpo.gc.ca>, <Phil.Greyson@dfo-mpo.gc.ca>, and <Tana.Worcester@dfo-mpo.gc.ca>

### **Acknowledgements**

<Mike.McMahon@dfo-mpo.gc.ca>! for providing code and advice in the use of `Mar.datawrangling`: A suite of tools for extracting, filtering and aggregating data from the Maritimes fisheries science databases <https://github.com/Maritimes/Mar.datawrangling>

The Atlantic R Working Group <https://github.com/AtlanticR>; in particular: Clark Richards and Brendan Wringe.  

