## **INFORMATION FROM NATIONAL AQUATIC SPECIES AT RISK PROGRAM**    

This section...

### **Species At Risk distribution and Critical Habitat data**    

#### ***DFO Species At Risk distribution (range)***      
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from Open Data  
Link: <https://open.canada.ca/data/en/dataset/e0fabad5-9379-4077-87b9-5705f28c490b>   
Quality Tier: High  
Security level: none  

#### ***Critical Habitat of Species At Risk***    
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from Open Data  
Link: <https://open.canada.ca/data/en/dataset/db177a8c-5d7d-49eb-8290-31e6a45d786c>   
Quality Tier: High  
Security level: none  

#### ***Area-specific SAR distribution and critical habitat search results***    

```{r echo=FALSE, results='asis'}
dist_table<-table_dist(sardist_sf,studyArea)
```
*
```{r dist result, comment="", prompt=TRUE, echo=FALSE, results='asis'}
Dist_overlap_result<-nrow(dist_table)
Query_output_dist<-if(Dist_overlap_result < 1){
  "The search area does not overlap with Species at Risk distribution range."
} else {
  "The search area overlaps with Species at Risk distribution range."
}
writeLines(Query_output_dist)
```

```{r echo=FALSE, results='asis'}
crit_table<-table_crit(ClippedCritHab_sf,studyArea, leatherback_sf)
```
*
```{r crit result, comment="", prompt=TRUE, echo=FALSE, results='asis'}
crit_result<-as.numeric(nrow(crit_table))
Query_output_crit<-if(crit_result < 1){
  "The search area does not overlap with defined Species at Risk Critical Habitat."
} else {
  "The search area overlaps with defined Species at Risk Critical Habitat."
}
writeLines(Query_output_crit)
```

```{r echo=FALSE, results='asis'}
sardist_wide<-plot_sardist(sardist_sf, studyArea, land10m_sf, 500)
sardist_zoom<-plot_sardist(sardist_sf, studyArea, land10m_sf, 50)
```

```{r, fig.height=6, fig.width=11, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with distribution (range) of Species At Risk listed as Endangered, Threatened or Special Concern under the <i>Species At Risk Act</i>. This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.'}
grid.arrange(sardist_wide, sardist_zoom,
             bottom = expression(paste("Longitude ",degree,"N",sep="")),
             left = expression(paste("Latitude ",degree,"N",sep="")),
             nrow = 1,
             widths = c(2.5/5, 2.5/5))
```

<br><br>

```{r echo=FALSE, results='asis'}
Dist_query_output_table<-if(Dist_overlap_result < 1){
  ""
} else {
  kable(dist_table[!duplicated(dist_table),], align="l", caption= "\\label{tab:dist_table}Quality Tier: High. Security level: none. Species At Risk listed as Endangered, Threatened or Special Concern under the <i>Species At Risk Act</i> for which the search polygon overlaps with their distribution (range). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create anthe authoritative SARA report.", booktabs = T, escape = F, linesep="")%>%
    kable_styling(bootstrap_options = "striped", full_width = T, position = "left")
}
if(Dist_overlap_result > 1){Dist_query_output_table}
```

<br><br>

```{r echo=FALSE, results='asis'}
crithab_wide<-plot_crithab(ClippedCritHab_sf, studyArea, land10m_sf, 500)
crithab_zoom<-plot_crithab(ClippedCritHab_sf, studyArea, land10m_sf, 50)
```

```{r, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with critical habitat of Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA). Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative report.', fig.height=6, fig.width=11}
grid.arrange(crithab_wide, crithab_zoom,
             bottom = expression(paste("Longitude ",degree,"N",sep="")),
             left = expression(paste("Latitude ",degree,"N",sep="")),
             nrow = 1,
             widths = c(2.5/5, 2.5/5))
```

<br><br>

```{r echo=FALSE, results='asis'}
Crithab_query_output_table<-if(crit_result < 1){
  ""
} else {
  kable(crit_table, caption="Quality Tier: High. Security level: none. Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA) for which the search polygon overlaps with their critical habitat. Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is <b>not</b> the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.", booktabs = T, escape = F, linesep="") %>%
    kable_styling(bootstrap_options = "striped", full_width = T, position = "left")}
if(crit_result > 0){Crithab_query_output_table}
```

