## **INFORMATION FROM NATIONAL AQUATIC SPECIES AT RISK PROGRAM**    

This section...

```{r Set up paths for National Aquatic Species at Risk Section, include=FALSE, cache=FALSE}
#fl <- list.files(dataPath,".csv")
ClippedCritHab <- st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/clipped_layers/ClipCritHab.shp", quiet=TRUE)
sardist<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/clipped_layers/sardist_4326.shp", quiet=TRUE)
leatherback_shp<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/LeatherBackTurtleCriticalHabitat/LBT_CH_2013.shp", quiet=TRUE)
```

### **Species At Risk distribution and Critical Habitat data**    

#### ***DFO Species At Risk distribution (range)*** {-}    
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from [Open Data](https://open.canada.ca/data/en/dataset/e0fabad5-9379-4077-87b9-5705f28c490b)  
Quality Tier: High  
Security level: none  

#### ***Critical Habitat of Species At Risk***    
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from [Open Data](https://open.canada.ca/data/en/dataset/db177a8c-5d7d-49eb-8290-31e6a45d786c)  
Quality Tier: High  
Security level: none  

#### ***Area-specific SAR distribution and critical habitat search results***    

*
```{r echo=FALSE, results='asis'}
intersect_dist <- st_intersection(sardist,PEZ_poly_sf)
intersect_dist$Common_Nam[intersect_dist$Common_Nam == "Sowerby`s Beaked Whale"] <- "Sowerby's Beaked Whale"
Common_Name<-intersect_dist$Common_Nam
Scientific_Name<-intersect_dist$Scientific
Population<-intersect_dist$Population
Area<-intersect_dist$Waterbody
intersect_dist_df<-as.data.frame(cbind(Common_Name, Scientific_Name, Population, Area))

#append columns from listed species 
dist_table<-merge(intersect_dist_df, listed_species, by='Scientific_Name')
dist_table<-dist_table %>% 
  transmute(Common_Name, Scientific_Name, Population, Area, Schedule.status, COSEWIC.status, Wild_Species)
dist_table<- dist_table %>% rename("SARA status"=Schedule.status,
                                   "COSEWIC listing"=COSEWIC.status,
                                   "Wild Species listing"=Wild_Species,
                                   "Common Name"=Common_Name,
                                   "Scientific Name"=Scientific_Name)
Dist_overlap_result<-length(dist_table)
Query_output_dist<-if(Dist_overlap_result < 1){
  "The search area does not overlap with Species at Risk distribution range."
} else {
  "The search area overlaps with Species at Risk distribution range."
}
Query_output_dist<-noquote(Query_output_dist)
writeLines(Query_output_dist)
```

```{r echo=FALSE, results='asis'}
sardist_wide<-plot_sardist(sardist, PEZ_poly_sf, land_sf, 400)
sardist_zoom<-plot_sardist(sardist, PEZ_poly_sf, land_sf, 50)
```

```{r, fig.height=8, fig.width=11, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with distribution (range) of Species At Risk listed as Endangered, Threatened or Special Concern under the <i>Species At Risk Act</i>. This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.'}
grid.arrange(sardist_wide, sardist_zoom,
             bottom = expression(paste("Longitude ",degree,"N",sep="")),
             left = expression(paste("Latitude ",degree,"N",sep="")),
             nrow = 1)
```


```{r echo=FALSE, results='asis'}
Dist_query_output_table<-if(Dist_overlap_result < 1){
  ""
} else {
  kable(dist_table[!duplicated(dist_table),], align="l", caption= "\\label{tab:dist_table}Quality Tier: High. Security level: none. Species At Risk listed as Endangered, Threatened or Special Concern under the <i>Species At Risk Act</i> for which the search polygon overlaps with their distribution (range). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create anthe authoritative SARA report.", booktabs = T, escape = F, linesep="")%>%
    kable_styling(bootstrap_options = "striped", full_width = T, position = "left")
}
if(Dist_overlap_result > 1){Dist_query_output_table}
```

*
```{r echo=FALSE, results='asis'}
crit_intersect <- st_intersection(ClippedCritHab,PEZ_poly_sf)
crit_table<-data.frame(CommonName=crit_intersect$Common_Nam,
                       Population=crit_intersect$Population, 
                       Area=crit_intersect$Waterbody,
                       SARA_status=crit_intersect$SARA_Statu)

leatherback_intersect <- st_intersection(leatherback_shp,PEZ_poly_sf)
leatherback_result<-as.numeric(nrow(leatherback_intersect))
leatherback_table<-data.frame(CommonName="",Population="", Area="", SARA_status="")
leatherback_table[1,1]<-"Leatherback Sea Turtle"
leatherback_table[1,2]<-NA
leatherback_table[1,3]<-leatherback_intersect$AreaName
leatherback_table[1,4]<-"Endangered"

crit_table2<-bind_rows(crit_table,leatherback_table)

crithab_result<-as.numeric(nrow(crit_table2))
Query_output_crit<-if(crithab_result < 1){
  "The search area does not overlap with defined Species at Risk Critical Habitat."
} else {
  "The search area overlaps with defined Species at Risk Critical Habitat."
}
Query_output_crit<-noquote(Query_output_crit)
writeLines(Query_output_crit)
```


```{r echo=FALSE, results='asis'}
Crithab_query_output_table<-if(crithab_result < 1){
  ""
} else {
  kable(crit_table2, caption="Quality Tier: High. Security level: none. Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA) for which the search polygon overlaps with their critical habitat. Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is <b>not</b> the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.", booktabs = T, escape = F, linesep="") %>%
    kable_styling(bootstrap_options = "striped", full_width = T, position = "left")}
if(crithab_result > 0){Crithab_query_output_table}
```

```{r echo=FALSE, results='asis'}
crithab_wide<-plot_crithab(ClippedCritHab, PEZ_poly_sf, land_sf, 600)
crithab_zoom<-plot_crithab(ClippedCritHab, PEZ_poly_sf, land_sf, 50)
```

```{r, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with critical habitat of Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA). Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative report.', fig.height=8, fig.width=11}
grid.arrange(crithab_wide, crithab_zoom,
             bottom = expression(paste("Longitude ",degree,"N",sep="")),
             left = expression(paste("Latitude ",degree,"N",sep="")),
             nrow = 1)
```
