```{r load packages, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(easypackages)
libraries("Mar.datawrangling","knitr","kableExtra","rgdal","maps","lubridate","raster","RCurl","sf","stringr","ggplot2","data.table","gridExtra","dplyr",
          "stars","ggspatial","tidyverse","standardPrintOutput")
```
```{r load source functions, include=FALSE, cache=FALSE}
source("site_map.r")
source("site_map_new.r")
source("filter_wsdb.r")
source("filter_narwc_new.r")
source("EBSA.R")
source("leatherback.R")
source("cetacean_priority_areas.R")
source("Blue_Whale_habitat.R")
source("fn_SearchRVData.r")
source("fn_Search_ISDB_Data.r")
source("fn_maps.r")
source("fn_Search_MARFIS_Data.r")
```
```{r Define Site, include=FALSE, cache=FALSE}
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
minYear <- 2002
```
```{r set up paths, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
land_sf<-st_read("../../../Data/Boundaries/Coast50K/Coastline50k_SHP/Land_AtlCanada_ESeaboardUS.shp", quiet=TRUE)
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
PEZ_poly_st<-st_as_sf(PEZ_poly)
```
```{r load SAR species lists, include=FALSE, cache=FALSE}
listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
cetacean_list<-c("Beluga Whale", "Humpback Whale" , "North Atlantic Right Whale", "Fin Whale", "Northern Bottlenose Whale", "Harbour Porpoise", "Killer Whale", "Blue Whale", "Sei Whale", "Sowerby's Beaked Whale")
other_species_list<-c("Loggerhead Sea Turtle", "Atlantic Walrus", "Harbour Seal Lacs des Loups Marins subspecies", "Leatherback Sea Turtle")
listed_cetacean_species<-subset(listed_species, Common_Name %in% cetacean_list)
listed_other_species<-subset(listed_species, Common_Name %in% other_species_list)
listed_fish_invert_species<-listed_species[ ! listed_species$Common_Name %in% c(other_species_list,cetacean_list), ]
```

## **Habitat Information**

### **Intertidal Vegetation and Rockweed**
Satellite-based Maps of Intertidal Vegetation and Rockweed presence polygons
Contact: <Gordana.Lazin@dfo-mpo.gc.ca>
Last retrieved on: February 17 2021 
Quality Tier: **Medium**
Security level: none.

### **Area-specific Intertidal Vegetation and Rockweed search results**  

"There are no relevant records of predicted intertidal vegetation for this search area" OR "There are There are relevant records of predicted intertidal vegetation for this search area."  

Figure x. Quality Tier: Medium. Security level: none. Map showing the search area defined by the user: location of the proposed project/activity is highlighted in yellow (input polygon), surrounded by an exposure zone or user-defined buffer area in blue. Figure shows overlap between the search are and satellite-based maps of intertidal vegetation and rockweed presence polygons. 