---
title: "Reproducible Report for Species in Maritimes Region - for Internal DFO use only"
author: Synthesis prepared by the Reproducible Reporting Team, steering committee and advisors.
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---
```{r Clean up environment for new report, echo=FALSE, include=FALSE, cache=FALSE}
rm(list=ls(all=TRUE))
```
```{r setup, echo=FALSE, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(Mar.datawrangling)
library(Mar.utils)
library(knitr)
library(kableExtra)
library(rgdal)
library(maps)
library(lubridate)
library(raster)
library(RCurl)
library(sf)
library(stringr)
library(ggplot2)
library(data.table)
library(gridExtra)
library(dplyr)
library(stars)
library(ggspatial)
#install.packages("remotes")
#remotes::install_github("terminological/standard-print-output")
#library(standardPrintOutput)
#These next two lines are necessary for the generation of the water mark on all plots.
#install.packages("devtools")
#devtools::install_github("JosephCrispell/basicPlotteR")
library(basicPlotteR)
```
```{r source functions, echo=FALSE, include=FALSE, cache=FALSE}
# the .Rmd file sets the working directory based on where it resides and ignores the wd set by the .RProj file
source("site_map.r")
source("site_map_new.r")
source("plot_wsdb.r")
source("filter_wsdb.r")
source("plot_narwc.r")
source("plot_narwc.r")
source("filter_narwc_new.r")
# ## Another option is to load directly from the web and read in using Rcurl
#function_urls <- c("https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/site_map.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/plot_wsdb.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/filter_wsdb.r")
#for (i in function_urls){eval(parse(text = RCurl::getURL(i, ssl.verifypeer = FALSE)),envir=.GlobalEnv)}
source("EBSA.R")
source("plot_EBSA.R")
source("leatherback.R")
source("plot_leatherback.R")
source("plot_obis.R")
source("plot_cws.R")
source("plot_crithab.R")
source("plot_crithab_zoom.R")
source("plot_inat.R")
source("cetacean_priority_areas.R")
source("plot_cetaceans_4grid.R")
source("Blue_Whale_habitat.R")
source("plot_bw_hab.R")
source("plot_bw_hab_zoom.R")
```
```{r Define Site, echo=FALSE, include=FALSE, cache=FALSE}
#HTML output of code is automatically generated in Catalina's computer: RProjects/MSP/SearchPEZ/code
#output of get.data.R should be manually moved to RProjects/MSP/SearchPEZ/outputs/sitenamexkm  
#AquaSiteName <- "TestGully"
#PEZversion <- "Polygon"
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
UserComments <- "Example - First search effort for suitability of a site for a project planned for 2025. Project is expected to have positive/negative impacts on surrounding area. It is likely that some kind of monitoring will be required during operations. Make sure to follow up on all sections pertaining to a particular taxonomic group as this seems to be a concern of the stakeholders."
#make sure you delete all shp.xml files in the folder to be able to run this code
minYear <- 1970
maxYear <- 2020
```
```{r Set up paths, echo=FALSE, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
#create a folder in outputs with the name of the site and buffer area search
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
# setwd(dataPath) #not needed?
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
opendataPath = "../../../Data/mar.wrangling/RVSurvey_OpenData"
#wsdbFile <- file.path("../inputs","WSDB_GLazin_AllSpecies_2018.csv")
wsdbFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/WSDB","MarWSDBSightingsForCGomez_27Oct2020.csv")
narwcFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/NARWC","NARWC_09-18-2020.csv")
```
```{r read baseline data for maps, echo=FALSE, include=FALSE, cache=FALSE}
landGDB <- "../../../Data/Boundaries/Coast50K/Coastline50k.gdb"
land <- readOGR(dsn=landGDB,layer="Land50k_Maritimes",stringsAsFactors=F)
land_sf <- st_as_sf(land)
```
```{r Define PEZ!, echo=FALSE, include=FALSE, cache=FALSE}
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
#PEZ_poly_st<-st_read(../../../Data/Zones/SearchPEZpolygons/PEZ_TestGullyPolygon.shp", quiet=TRUE)
PEZ_poly_st<-st_as_sf(PEZ_poly)
```
```{r Species listed by Wildspecies, COSEWIC and SARA in the Maritime Region, echo=FALSE, include=FALSE, cache=FALSE}

listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
MARSAR_com <- listed_species$Common_Name
MARSAR_com_up <- listed_species$Common_Name_upper
MARSAR_sci <- listed_species$Scientific_Name
MARSAR_sci_up <- listed_species$Scientific_Name_upper
```

# **Search Results for Species at Risk**  

## **Sea Turtle Section**
```{r Set up paths for leatherback turtle files, echo=FALSE, include=FALSE, cache=FALSE}
leatherback_shp <- st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/LeatherBackTurtleCriticalHabitat/LBT_CH_2013.shp", quiet=TRUE)
```

Data products used in this section:

### **Leatherback Sea Turtle Important Habitat**  
Contact:  
Tier of confidence: **High**  
Search year: 
Disclaimers, caveats, sources of uncertainty  

#### **Draft: Leatherback Search Results**  

```{r echo=FALSE, results='asis'}
#function for overlap
leatherback_overlap(leatherback_shp=leatherback_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for sector
leatherback_area(leatherback_shp=leatherback_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r Fig X PEZ polygon with all maritime leatherback polygons, fig.height=8, fig.width=11}
plot_leatherback(leatherback_shp, PEZ_poly_st, land_sf)
```
**Fig. 7:** Map showing Leatherback Sea Turtle Important Habitat (green) relative to the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow).
