---
title: "Reproducible Report for Species in Maritimes Region - for Internal DFO use only"
author: Synthesis prepared by the Reproducible Reporting Team, steering committee and advisors.
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---
```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(Mar.datawrangling)
library(Mar.utils)
library(knitr)
library(kableExtra)
library(rgdal)
library(maps)
library(lubridate)
library(raster)
library(RCurl)
library(sf)
library(stringr)
library(ggplot2)
library(data.table)
library(gridExtra)
library(dplyr)
library(stars)
library(ggspatial)
library(tidyverse)
#These next two lines are necessary for the generation of the water mark on all plots.
#install.packages("remotes")
#remotes::install_github("terminological/standard-print-output")
library(standardPrintOutput)
```
```{r source functions, include=FALSE, cache=FALSE}
# the .Rmd file sets the working directory based on where it resides and ignores the wd set by the .RProj file
source("site_map.r")
source("site_map_new.r")
source("filter_wsdb.r")
source("filter_narwc_new.r")
# ## Another option is to load directly from the web and read in using Rcurl
#function_urls <- c("https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/site_map.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/plot_wsdb.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/filter_wsdb.r")
#for (i in function_urls){eval(parse(text = RCurl::getURL(i, ssl.verifypeer = FALSE)),envir=.GlobalEnv)}
source("EBSA.R")
source("leatherback.R")
source("cetacean_priority_areas.R")
source("Blue_Whale_habitat.R")
source("fn_SearchRVData.r")
source("fn_maps.r")
```
```{r Define Site, include=FALSE, cache=FALSE}
#HTML output of code is automatically generated in Catalina's computer: RProjects/MSP/SearchPEZ/code
#output of get.data.R should be manually moved to RProjects/MSP/SearchPEZ/outputs/sitenamexkm  
#AquaSiteName <- "TestGully"
#PEZversion <- "Polygon"
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
UserComments <- "Example - First search effort for suitability of a site for a project planned for 2025. Project is expected to have positive/negative impacts on surrounding area. It is likely that some kind of monitoring will be required during operations. Make sure to follow up on all sections pertaining to a particular taxonomic group as this seems to be a concern of the stakeholders."
#make sure you delete all shp.xml files in the folder to be able to run this code
minYear <- 1970
maxYear <- 2020
```
```{r Set up paths, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
#create a folder in outputs with the name of the site and buffer area search
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
# setwd(dataPath) #not needed?
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
#wsdbFile <- file.path("../inputs","WSDB_GLazin_AllSpecies_2018.csv")
wsdbFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/WSDB","MarWSDBSightingsForCGomez_27Oct2020.csv")
narwcFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/NARWC","NARWC_09-18-2020.csv")
```
```{r read baseline data for maps, include=FALSE, cache=FALSE}
landGDB <- "../../../Data/Boundaries/Coast50K/Coastline50k.gdb"
land <- readOGR(dsn=landGDB,layer="Land50k_Maritimes",stringsAsFactors=F)
land_sf <- st_as_sf(land)
```
```{r Define PEZ!, include=FALSE, cache=FALSE}
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
#PEZ_poly_st<-st_read(../../../Data/Zones/SearchPEZpolygons/PEZ_TestGullyPolygon.shp", quiet=TRUE)
PEZ_poly_st<-st_as_sf(PEZ_poly)
```
```{r Species listed by Wildspecies, COSEWIC and SARA in the Maritime Region, include=FALSE, cache=FALSE}

listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
MARSAR_com <- listed_species$Common_Name
MARSAR_com_up <- listed_species$Common_Name_upper
MARSAR_sci <- listed_species$Scientific_Name
MARSAR_sci_up <- listed_species$Scientific_Name_upper
```

# **Search Results for Species at Risk**  

## **Ecologically and Biologically Significant Areas (EBSA)**  
Contact: <https://open.canada.ca/data/en/dataset/d2d6057f-d7c4-45d9-9fd9-0a58370577e0>  
Tier of confidence: **High**  
Search year:  

Ecologically and Biologically Significant Areas (EBSAs) are areas within Canada's oceans that have been identified through formal scientific assessments as having special biological or ecological significance when compared with the surrounding marine ecosystem.  

Failure to define an area as an EBSA does not mean that it is unimportant ecologically. All areas serve ecological functions to some extent and require sustainable management. Rather, areas identified as EBSAs should be viewed as the most important areas where, with existing knowledge, regulators and marine users should be particularly risk averse to ensure ecosystems remain healthy and productive.

#### **EBSA search results**
```{r read other species database files for other species, include=FALSE, cache=FALSE}
EBSA <- st_read("../../../Data/Zones/DFO_EBSA/DFO_EBSA.shp", quiet=TRUE)
EBSA_shp <- st_transform(EBSA, crs = 4326)
```

```{r echo=FALSE, results='asis'}
#function for overlap
EBSA_overlap(EBSA_shp=EBSA_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for report
EBSA_report(EBSA_shp=EBSA_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for report url
EBSA_reporturl(EBSA_shp=EBSA_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for location
EBSA_location(EBSA_shp=EBSA_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for bioregion
EBSA_bioregion(EBSA_shp=EBSA_shp, PEZ_poly_st= PEZ_poly_st)
```

```{r Fig X PEZ polygon with all maritime EBSA polygons, fig.height=8, fig.width=11, fig.cap="Map showing Ecologically and Biologically Significant Areas (EBSA; pink) relative to the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow)."}
plot_EBSA(EBSA_shp, PEZ_poly_st, land_sf)
```