## **INFORMATION FROM NATIONAL AQUATIC SPECIES AT RISK PROGRAM**    

This section...

```{r Set up paths for National Aquatic Species at Risk Section, include=FALSE, cache=FALSE}
fl <- list.files(dataPath,".csv")
ClippedCritHab <- st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/clipped_layers/ClipCritHab.shp", quiet=TRUE)
sardist<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/clipped_layers/sardist_4326.shp", quiet=TRUE)
leatherback_shp<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/LeatherBackTurtleCriticalHabitat/LBT_CH_2013.shp", quiet=TRUE)
```

### **Species At Risk distribution and Critical Habitat data**    

#### ***DFO Species At Risk distribution (range)*** {-}    
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from [Open Data](https://open.canada.ca/data/en/dataset/e0fabad5-9379-4077-87b9-5705f28c490b)  
Quality Tier: High  
Security level: none  

#### ***Critical Habitat of Species At Risk***    
Contact: <info@dfo-mpo.gc.ca>  
Last retrieved on: October 1 2020 from [Open Data](https://open.canada.ca/data/en/dataset/db177a8c-5d7d-49eb-8290-31e6a45d786c)  
Quality Tier: High  
Security level: none  

#### ***Area-specific SAR distribution and critical habitat search results***    

*
```{r echo=FALSE, results='asis'}
intersect_dist <- st_intersection(sardist,PEZ_poly_st)
intersect_dist$Common_Nam[intersect_dist$Common_Nam == "Sowerby`s Beaked Whale"] <- "Sowerby's Beaked Whale"
Common_Name<-intersect_dist$Common_Nam
Scientific_Name<-intersect_dist$Scientific
Population<-intersect_dist$Population
Area<-intersect_dist$Waterbody
intersect_dist_df<-as.data.frame(cbind(Common_Name, Scientific_Name, Population, Area))

#append columns from listed species 
dist_table<-merge(intersect_dist_df, listed_species, by='Scientific_Name')
dist_table<-dist_table %>% 
  transmute(Common_Name, Scientific_Name, Population, Area, Schedule.status, COSEWIC.status, Wild_Species)
dist_table<- dist_table %>% rename("SARA status"=Schedule.status,
                                   "COSEWIC listing"=COSEWIC.status,
                                   "Wild Species listing"=Wild_Species,
                                   "Common Name"=Common_Name,
                                   "Scientific Name"=Scientific_Name)
Dist_overlap_result<-length(dist_table)
Query_output_dist<-if(Dist_overlap_result < 1){
  "The search area does not overlap with Species at Risk distribution range."
} else {
  "The search area overlaps with Species at Risk distribution range."
}
Query_output_dist<-noquote(Query_output_dist)
writeLines(Query_output_dist)
```
```{r, fig.height=8, fig.width=11, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with distribution (range) of Species At Risk listed as Endangered, Threatened or Special Concern under the Species At Risk Act. This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.'}
plot_sardist(sardist, PEZ_poly_st, land_sf)
```

```{r, fig.height=8, fig.width=11, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with distribution (range) of Species At Risk listed as Endangered, Threatened or Special Concern under the Species At Risk Act. This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.'}
plot_sardist_zoom(sardist, PEZ_poly_st, land_sf)
```


```{r echo=FALSE, results='asis'}
Dist_query_output_table<-if(Dist_overlap_result < 1){
  ""
} else {
  kable(dist_table[!duplicated(dist_table),], align="l", caption= "\\label{tab:dist_table}Quality Tier: High. Security level: none. Species At Risk listed as Endangered, Threatened or Special Concern under the Species At Risk Act for which the search polygon overlaps with their distribution (range). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create anthe authoritative SARA report.", booktabs = T, escape = F, linesep="")%>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
if(Dist_overlap_result > 1){Dist_query_output_table}
```

*
```{r echo=FALSE, results='asis'}
crit_intersect <- st_intersection(ClippedCritHab,PEZ_poly_st)
CommonName<-crit_intersect$Common_Nam
Population<-crit_intersect$Population
Area<-crit_intersect$Waterbody
SARA_status<-crit_intersect$SARA_Statu
crit_table<-data.table(expand.grid(CommonName = CommonName, Population = Population, Area=Area, SARA_status=SARA_status))
crithab_result<-as.numeric(nrow(crit_table))
Query_output_crit<-if(crithab_result < 1){
  "The search area does not overlap with defined Species at Risk Critical Habitat."
} else {
  "The search area overlaps with defined Species at Risk Critical Habitat."
}
Query_output_crit<-noquote(Query_output_crit)
writeLines(Query_output_crit)
```


```{r echo=FALSE, results='asis'}
Crithab_query_output_table<-if(crithab_result < 1){
  ""
} else {
  kable(crit_table[!duplicated(crit_table),], caption="\\label{tab:crit_table}Quality Tier: High. Security level: none. Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA) for which the search polygon overlaps with their critical habitat. Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is <b>not</b> the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative SARA report.", booktabs = T, escape = F, linesep="") %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
if(crithab_result > 1){print(Crithab_query_output_table)}
```


```{r, fig.cap='Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with critical habitat of Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA). Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the [SAR Public Registry](http://www.sararegistry.gc.ca). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk [GIS tool](http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US) to create an authoritative report.', fig.height=8, fig.width=11}
plot_crithab_zoom<-plot_crithab_zoom(ClippedCritHab, PEZ_poly_st, land_sf)
```
