---
title: "Reproducible Report for Species in Maritimes Region - for Internal DFO use only"
author: Synthesis prepared by the Reproducible Reporting Team, steering committee and advisors.
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---
```{r Clean up environment for new report, echo=FALSE, include=FALSE, cache=FALSE}
rm(list=ls(all=TRUE))
```
```{r setup, echo=FALSE, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(Mar.datawrangling)
library(Mar.utils)
library(knitr)
library(kableExtra)
library(rgdal)
library(maps)
library(lubridate)
library(raster)
library(RCurl)
library(sf)
library(stringr)
library(ggplot2)
library(data.table)
library(gridExtra)
library(dplyr)
library(stars)
library(ggspatial)
#install.packages("remotes")
#remotes::install_github("terminological/standard-print-output")
#library(standardPrintOutput)
#These next two lines are necessary for the generation of the water mark on all plots.
#install.packages("devtools")
#devtools::install_github("JosephCrispell/basicPlotteR")
library(basicPlotteR)
```
```{r source functions, echo=FALSE, include=FALSE, cache=FALSE}
# the .Rmd file sets the working directory based on where it resides and ignores the wd set by the .RProj file
source("site_map.r")
source("site_map_new.r")
source("plot_wsdb.r")
source("filter_wsdb.r")
source("plot_narwc.r")
source("plot_narwc.r")
source("filter_narwc_new.r")
# ## Another option is to load directly from the web and read in using Rcurl
#function_urls <- c("https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/site_map.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/plot_wsdb.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/filter_wsdb.r")
#for (i in function_urls){eval(parse(text = RCurl::getURL(i, ssl.verifypeer = FALSE)),envir=.GlobalEnv)}
source("EBSA.R")
source("plot_EBSA.R")
source("leatherback.R")
source("plot_leatherback.R")
source("plot_obis.R")
source("plot_cws.R")
source("plot_crithab.R")
source("plot_crithab_zoom.R")
source("plot_inat.R")
source("cetacean_priority_areas.R")
source("plot_cetaceans_4grid.R")
source("Blue_Whale_habitat.R")
source("plot_bw_hab.R")
source("plot_bw_hab_zoom.R")
```
```{r Define Site, echo=FALSE, include=FALSE, cache=FALSE}
#HTML output of code is automatically generated in Catalina's computer: RProjects/MSP/SearchPEZ/code
#output of get.data.R should be manually moved to RProjects/MSP/SearchPEZ/outputs/sitenamexkm  
#AquaSiteName <- "TestGully"
#PEZversion <- "Polygon"
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
UserComments <- "Example - First search effort for suitability of a site for a project planned for 2025. Project is expected to have positive/negative impacts on surrounding area. It is likely that some kind of monitoring will be required during operations. Make sure to follow up on all sections pertaining to a particular taxonomic group as this seems to be a concern of the stakeholders."
#make sure you delete all shp.xml files in the folder to be able to run this code
minYear <- 1970
maxYear <- 2020
```
```{r Set up paths, echo=FALSE, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
#create a folder in outputs with the name of the site and buffer area search
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
# setwd(dataPath) #not needed?
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
opendataPath = "../../../Data/mar.wrangling/RVSurvey_OpenData"
#wsdbFile <- file.path("../inputs","WSDB_GLazin_AllSpecies_2018.csv")
wsdbFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/WSDB","MarWSDBSightingsForCGomez_27Oct2020.csv")
narwcFile <- file.path("../../../Data/NaturalResources/Species/Cetaceans/NARWC","NARWC_09-18-2020.csv")
```
```{r read baseline data for maps, echo=FALSE, include=FALSE, cache=FALSE}
landGDB <- "../../../Data/Boundaries/Coast50K/Coastline50k.gdb"
land <- readOGR(dsn=landGDB,layer="Land50k_Maritimes",stringsAsFactors=F)
land_sf <- st_as_sf(land)
```
```{r Define PEZ!, echo=FALSE, include=FALSE, cache=FALSE}
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
#PEZ_poly_st<-st_read(../../../Data/Zones/SearchPEZpolygons/PEZ_TestGullyPolygon.shp", quiet=TRUE)
PEZ_poly_st<-st_as_sf(PEZ_poly)
```
```{r Species listed by Wildspecies, COSEWIC and SARA in the Maritime Region, echo=FALSE, include=FALSE, cache=FALSE}

listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
MARSAR_com <- listed_species$Common_Name
MARSAR_com_up <- listed_species$Common_Name_upper
MARSAR_sci <- listed_species$Scientific_Name
MARSAR_sci_up <- listed_species$Scientific_Name_upper
```

# **Search Results for Species at Risk**  

## **Cetacean Section**

```{r Set up paths for Cetacean section, echo=FALSE, include=FALSE, cache=FALSE}
#read narwc file
narwc <- read.csv(narwcFile)
narwc <- narwc[which(narwc$YEAR>=minYear),]
#read wsdb file
wsdb <- read.csv(wsdbFile)
wsdb <- wsdb[which(wsdb$YEAR>=minYear),]
#read obis file
obis<-read.csv("../../../Data/NaturalResources/Species/OBIS_CBIF_iNaturalist/OBIS_MAR_priority_records.csv")
obis<-subset(obis,year>minYear,year<maxYear)
#read cws file
cws<-read.csv("../../../Data/NaturalResources/Species/CWS_ECCC/CWS_ECCC_OBIS_records.csv")
cws<-subset(cws,year>minYear,year<maxYear)
#read iNaturalist file
inat<-read.csv("../../../Data/NaturalResources/Species/OBIS_CBIF_iNaturalist/iNaturalist_MAR_priority_records.csv")
inat<-subset(inat,datetime>minYear,datetime<maxYear)
#convert fin whale raster to sf object
fin_whale<- raster("../../../Data/NaturalResources/Species/Cetaceans/PriorityAreas_FGP/Fin_Whale.tif")
fin_whale[fin_whale==0] <- NA
fin_whale_sf<-st_as_stars(fin_whale)%>%st_as_sf()
#convert harbour porpoise raster to sf object
harbour_porpoise<- raster("../../../Data/NaturalResources/Species/Cetaceans/PriorityAreas_FGP/Harbour_Porpoise.tif")
harbour_porpoise[harbour_porpoise==0] <- NA
harbour_porpoise_sf<-st_as_stars(harbour_porpoise)%>%st_as_sf()
#convert humpback whale raster to sf object
humpback_whale<- raster("../../../Data/NaturalResources/Species/Cetaceans/PriorityAreas_FGP/Humpback_Whale.tif")
humpback_whale[humpback_whale==0] <- NA
humpback_whale_sf<-st_as_stars(humpback_whale)%>%st_as_sf()
#convert sei whale raster to sf object
sei_whale<- raster("../../../Data/NaturalResources/Species/Cetaceans/PriorityAreas_FGP/Sei_Whale.tif")
sei_whale[sei_whale==0] <- NA
sei_whale_sf<-st_as_stars(sei_whale)%>%st_as_sf()
Blue_32198 <- st_read("../../../Data/NaturalResources/Species/Cetaceans/BlueWhaleHabitat_FGP/BlueWhaleHabitat_HabitatBaleineBleue.shp", quiet=TRUE)
Blue_4326 <- st_transform(Blue_32198, crs = 4326)
Blue_4326b<-setNames(Blue_4326, replace(names(Blue_4326), names(Blue_4326) == 'activité', 'activite'))
```

This section summarizes information available from visual sightings and modeled predictions of suitable habitat to identify cetacean species in the search area. While numerous species have been identified in this region, only those listed by SARA, or assessed by COSEWIC and Wild species listings are summarized. Outputs include results from cetacean surveys conducted over previous years in both Canadian and American waters in the northwest Atlantic. However, the majority of cetacean sightings for the Scotian Shelf bioregion are opportunistic in nature, and comprehensive systematic surveys on the occurrence or distribution of cetaceans across the entire region are very limited. Results in this report can provide some information on distribution, but cannot be interpreted as comprehensive indices of abundance as they often do not cover the entirety of a species’ range. A future comprehensive analysis of any associated observational effort for these data is necessary to fully characterize cetacean distribution in the area over temporal and spatial scales. Consequently, outputs of this report indicate species presence only, and an absence of a species in an area should be interpreted as an absence of reporting not as an absence of the species in the area.

This section does not endeavor to describe every source of information provided below. Please refer to the publications listed below for additional information and please contact Angelia Vanderlaan (North Atlantic right whales) and Hilary Moors-Murphy (beaked and other baleen whales) to consider additional sources of information (e.g. acoustic detections) to support species-specific conclusions about habitat preferences and possible or known occurrence near and within the search area. This report does not provide the best available information for North Atlantic right whales; please refer to WHALEMAP for the latest North Atlantic right whale observations: https://whalemap.ocean.dal.ca/ Note that WhaleMap (Johnson 2018) reports primarily presence-only datasets and only reflects results for areas that are being monitored.

### **Sightings**

Data products used in this section: 

#### **North Atlantic Right Whale Consortium (NARWC) Database**  
Contact: <https://www.narwc.org/sightings-database.html>  
Last retrieved on: September 18 2020  
Tier of confidence: **High**  
Search year: 2000-2019 

The following information is summarized in the NARWC sightings database site <https://www.narwc.org/sightings-database.html>. The Sightings database contains records of thousands of sightings of right whales in the North Atlantic Ocean, as well as sightings of many other species of whales, dolphins, sea turtles, seals, and large fishes. It also contains survey data associated with many of these sightings that allow quantification of associated survey effort (note that the database does not include interpreted effort data as such). The sightings contained in the database come from a wide variety of contributors both Consortium members and others. Each record in the Sightings database represents a group of whales (i.e., a group of 3 whales has a single record just as a group of 1 does) and there may or may not be photographic proof of a given right whale sighting. The Sightings and Identification databases are periodically cross-referenced, so that individual identification data from the latter can be linked to sighting data from the former. For that reason, all sightings in the Identification database are eventually included in the Sightings database (with an approximate 1-year lag). Although the animals' identifications are not included in the Sightings database, the two databases can be linked on common fields. This review article of the Sightings Database provides potential users with important information.  

#### **The Whitehead Lab**  
Contact: <https://whiteheadlab.weebly.com/contact.html>  
Last retrieved on: January 12 2021  
Tier of confidence: **High**  
Search year:1988-2019  

The following information is summarized in this site (https://whiteheadlab.weebly.com/). The Whitehead Lab is based out of the Dalhousie University Department of Biology (Halifax, Nova Scotia). Since its inception in 1987, the Whitehead Lab has been home to over 75 postdoctoral, Ph.D., M.Sc., and undergraduate students. They primarily conduct conservation-based research on the behavior, ecology, and population biology of cetacean species (primarily, sperm whales, northern bottlenose whales, and long-finned pilot whales). The following is extracted from Whitehead (2013): data in this section is from incidental sightings of cetaceans during studies of northern bottlenose whales and, to a lesser extent, sperm whales in the northwest Atlantic, primarily in the Gully MPA, and adjacent canyons along the edge of the eastern Scotian Shelf (the Shortland Canyon and the Haldimand Canyon). Data collection is carried out in a comparable manner, and from two similar research vessels: a 10 m auxiliary sailing vessel, or a 13 m auxiliary sailing vessel. All groups of cetaceans sighted are recorded, together with time of sighting, species (where ascertainable), and position. As definitions of group size changed over the course of the study, sightings are very inaccurate and biased by factors such as the behaviour of the animals and the weather. Consequently, Whitehead (2013) only used group sighting data, rather than numbers of individuals. The Whitehead Lab provides cetacean sightings to the Whale Sightings Database (WSDB) at DFO. However, due to its high tier of confidence, this database is summarized independently in this report.

#### **Whale Sightings Database (WSDB)**  
Contact: XMARWhaleSightings@dfo-mpo.gc.ca  
Last retrieved on: October 27 2020  
Tier of confidence: **Low**  
Search year:  

The Marine Mammal and Pelagic Animals or Whale Sighting Database (WSDB) was implemented in 2002 by the Department of Fisheries and Oceans (DFO) to provide a central repository for opportunistic sightings of marine animals, especially cetaceans, and to improve accessibility to data from a variety of sources, including some data from marine mammal surveys and research activities conducted in the Bay of Fundy, on the Scotian Shelf (MacDonald et al. 2017), and more recently in the Gulf of St. Lawrence. Most sightings are collected on an opportunistic basis and observations come from individuals with a wide range of expertise and experience in marine mammal identification. Most data is gathered from platforms of opportunity that were vessel-based. For more information see <https://waves-vagues.dfo-mpo.gc.ca/Library/40642999.pdf>.

Note the following details about the data provided: 
-sightings with Behaviour coded as “dead”, “killed” or “stranded”  are not included.
-sightings with Animal Condition coded as "dead" are not included.
-sightings with Gear Impact recorded coded as  "Entangled - dead on gear" or "dead - not entangled"  are not included.

Please note the following important disclaimers for WSDB:
- Sighting effort has not been quantified (i.e., the data cannot be used to estimate true species density or abundance for an area).
- The sightings data are biased by when and where activities occur. Species absence in an area should be interpreted as an absence of reporting not as an absence of the species in the area.
- The sighting data have not been completely error-checked and may contain duplicate records.
- The quality of some of the sighting data is unknown. Most sightings are collected on an opportunistic basis and observations come from individuals with a wide range of expertise and experience in marine mammal identification. 
- There is significant uncertainty in the number of animals sighted per detection due to the challenges associated with quantifying the number of individuals observed in a group.
- The data represent an amalgamation of sightings across years and seasons. The sightings data are biased by when and where activities were conducted, thus, the sighting effort is not consistent across months, years or, areas. Therefore, apparent differences in seasonal or annual distribution  patterns in the sightings data should not be considered definitive. Unfavorable weather and reduced visual effort in winter, spring, and autumn likely account for the fewer sighting records in these seasons compared to summer.
- Many sightings which could not be identified to species are listed to the lowest taxonomic group possible.


#### **Ocean Biodiversity Information System (OBIS)**  
Contact: helpdesk@obis.org
Last retrieved on: January 27 2021 from OBIS  
Tier of confidence: **Medium**  
Search year: xxxxx  
```{r echo=FALSE, results='asis'}
cat("Time frame:", "insert results from min and max year search function for this db.")
```

OBIS is a global open-access data and information clearing-house on marine biodiversity for science, conservation and sustainable development. Their vision is to build and maintain a global alliance that collaborates with scientific communities to facilitate free and open access to, and application of, biodiversity and biogeographic data and information on marine life. OBIS search was conducted to find additional relevant records for listed by SARA, or assessed by COSEWIC and Wild species listing. A first level QC process was performed by removing duplicate reports that are also present in WSDB. Subsequent versions of this report will aim at expanding our QC efforts for this source of information. 

#### **CWS/Environment and Climate Change Canada**
Contact:  
Tier of confidence: **Low**  
```{r echo=FALSE, results='asis'}
cat("Time frame:", "insert results from min and max year search function for this db.")
```

#### **iNaturalist**
Contact:  
Tier of confidence: **Low**  
```{r echo=FALSE, results='asis'}
cat("Time frame:", "insert results from min and max year search function for this db.")
```
All data has been filtered by quality grade equal to "research".
For more information about iNaturalist, visit <https://www.inaturalist.org/>

### **Species Distribution Models**  

Data products used in this section:  

#### **Priority Areas to Enhance Monitoring of Cetaceans**  

Contact: Hilary.Moors-Murphy@dfo-mpo.gc.ca  
Site: <https://waves-vagues.dfo-mpo.gc.ca/Library/40869155.pdf>  
Last retrieved on: April 1 2020 from Open Data <https://open.canada.ca/data/en/dataset/c094782e-0d6f-4cc0-b5a3-58908493a433>  
Tier of confidence: **Low**  
Search year: 1975-2015  

The following text is verbatim from the Open data record. Species Distribution Models (SDM) were used to predict and identify priority areas for enhanced monitoring of cetaceans in eastern Canadian waters off Nova Scotia, Newfoundland and Labrador. This data set represents information presented in Gomez et al. (2020) and includes sighting records and SDM outputs for ten cetacean species with sufficient records (n > 450) and sightings only for an additional six species. For more information about sighting records including which were included in each SDM, please see Gomez et al. 2020. 
This study used a compilation of aerial- and vessel-based cetacean sightings data from 1975-2015 assembled in Gomez et al. (2017) from a variety of sources. Note that sightings data from many of these sources are not effort-corrected and apparent distribution patterns based on these opportunistic sightings data are biased by when and where survey activities were conducted. Unfavorable weather and reduced visual effort in winter, spring, and autumn likely account for the fewer sighting records in these seasons compared to summer. The dataset does not include dead animal, stranding, entanglement or entrapment data. While some of the databases include records obtained during the whaling period (catches or sightings recorded prior to 1975), for all analyses/modelling conducted in this study, only sightings of free-swimming whales obtained during the post-whaling period (1975-2015) were used. Quality control checks included discarding all records outside of our study area and removing redundant records (identical species, day, month, latitude and longitude).The data used do not reflect any updates or corrections to the databases that have occurred since the data were compiled in 2016. Sightings are not available for download here, please contact the original data sources listed below to obtain raw sightings data. 
This study represents an important initiative in eastern Canada to highlight key areas for cetacean monitoring in waters off Nova Scotia, Newfoundland and Labrador. Habitats with high suitability are interpreted as areas where cetacean monitoring efforts may be prioritized, and results can help direct future survey efforts. These model outputs used cetacean sightings from several decades and dynamic environmental predictors that used seasonal averages across multiple years. As proxies for prey availability, five predictor environmental variables were selected for the SDM: ocean depth, compound topographic index, sea surface temperature, areas of persistently high chlorophyll-a concentration, and regional chlorophyll-a magnitude. See Gomez et al. (2020) for further details on modelling methods. Persistent patterns over time (between 1975-2015) are the main patterns expected to be captured by these models. Further, SDM results presented here are not the same as species density maps; rather, they portray predicted suitable habitat based on environmental characteristics and sightings data that were not always derived from effort-based surveys. Consequently, the use of these models in marine spatial planning processes should be accompanied by complimentary approaches such as acoustic and visual validation of the SDM results as well as additional monitoring and modeling efforts. Please refer to Gomez et al. (2020) for examples on how to best use these data outputs. Future efforts will focus on using more recent data and improving these models to facilitate the inclusion of cetaceans in marine spatial planning processes that are currently underway. 


### **Important habitat: combined products used to inform critical habitat designation**  

Data products used in this section:  

#### **Important areas of Blue Whale feeding, foraging and migration in eastern Canada**  
Contact: <gddaiss-dmsaisb@dfo-mpo.gc.ca>  
Site: <https://waves-vagues.dfo-mpo.gc.ca/Library/40869155.pdf>  
Last retrieved on: October 30 2020 from Open Data <https://open.canada.ca/data/en/dataset/8fafd919-fcbe-43a3-a911-3d9461273441>  
Tier of confidence: **High**  

The following text is verbatim from the Open Data record. The distribution of blue whales (Balaenoptera musculus) is mainly dictated by key environmental drivers such as food requirements and availability, sea ice, and sometimes predation risks. It is common to find blue whales in upwelling regions, depending on topography (banks, slopes and shallow water) and thermal fronts, which are actually related to krill aggregations. Blue whales feed almost exclusively on euphausiids (krill) and have the highest absolute metabolic demands. A shapefile on the distribution of significant perennial areas of krill aggregation is available (DFO-ECCC). Important habitats for blue whales have therefore been identified by combining information on their distribution with krill aggregation areas (observed or predicted). In the waters of the St. Lawrence estuary, the presence of blue whales for foraging, feeding and migration has been observed throughout the year, with increased activity during the summer. It has been observed that feeding rates of blue whales are higher in shallow waters and during nocturnal periods (krill at surface), confirming that foraging near the surface is beneficial when possible. Their presence, their great fidelity to this important habitat for their activities and their behavior in the waters of the St. Lawrence in all seasons, show that blue whales are vulnerable to oil spills. They are vulnerable, not only indirectly through the reduction of prey and loss of suitable habitat, but also by their behavior requiring active activity on the surface and near the coast. The data used to create the polygons using the bounding box method comes from multiple sources of information: hunting records, photo-identification, coastal, airborne and boat surveys, passive acoustic monitoring, radio and satellite telemetry , ice trapping, anecdotal observations and modeling (please see a list of all sources of information used in Lesaqe et al. 2018 and the open data record <https://open.canada.ca/data/en/dataset/8fafd919-fcbe-43a3-a911-3d9461273441>).

#### **NARWC Search Results**
```{r read narwc and filter SARA & COSEWIC cetacean spp, echo=FALSE, include=FALSE, cache=FALSE}
narwc_filter <- filter_narwc_new(narwc)
narwc_filter <- narwc_filter %>% rename(Scientific_Name = SPECCODE)
narwc_filter <- merge(narwc_filter, listed_species, by='Scientific_Name')
```
```{r echo=FALSE, results='asis'}
narwc_plot_locations<-st_as_sf(narwc_filter, coords = c("LONGITUDE", "LATITUDE"), crs = 4326)
intersect_narwc <- st_intersection(narwc_plot_locations,PEZ_poly_st)
x<-as.numeric(nrow(intersect_narwc))
Common_Name<-intersect_narwc$Common_Name
Scientific_Name<-intersect_narwc$Scientific_Name
intersect_narwc_df<-as.data.frame(cbind(Common_Name, Scientific_Name, row.names = FALSE))

#append columns from listed species 
narwc_table<-merge(intersect_narwc_df, listed_species, by='Scientific_Name')
narwc_table2<-narwc_table %>% 
  transmute(Common_Name, Scientific_Name, Schedule.status, COSEWIC.status, Wild_Species)
narwc_table3<- narwc_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )

Query_output_narwc<-if(x < 1){
  "There are no records in the NARWC data base for this search area."
} else {
  "The NARWC database contains records of whales in this search area."
}

Query_output_narwc2<-noquote(Query_output_narwc)

writeLines(c(Query_output_narwc2), sep = "\n")
```

**Table 7:** Cetaceans with observations contained in the NARWC database within the search polygon area.

```{r echo=FALSE, results='asis'}
NARWC_query_output_table<-if(x < 1){
  ""
} else {
  kable(narwc_table3[!duplicated(narwc_table3),], row.names=F) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
  
}

NARWC_query_output_table

```
```{r Fig 5 NARWC Database, fig.height=8, fig.width=11}
par(mfrow = c(1,2), mar=c(0, 0, 5.5, 0) + 2)#it goes c(bottom, left, top, right)
narwc_plot<-site_map_new(PEZ_poly,PEZ_poly_st,site_sf,land_sf,5)+
  geom_point(data = narwc_filter, aes(x = LONGITUDE, y = LATITUDE, color=Common_Name), 
             size = 3.5)
print(narwc_plot + scale_colour_manual(values=c("darkgoldenrod1", "chartreuse4", "black", "#00AFBB","#B337A3", "#0827EF", "#EF6408", "#F5A4E7")))

```
**Fig. 4:** The tier of confidence of this map for internal use only is **High**. Map represents cetacean sightings from the North Atlantic Right Whale Consortium (NARWC) sightings database. Absence of a species in the map should be interpreted as an absence of reporting, not as an absence of the species in the area (e.g. field efforts were primarily conducted towards the Northeast US, sightings information underrepresents presence of cetaceans as this search does not include acoustic detections). For more information, please visit: https://www.narwc.org/sightings-database.html

#### **Whitehead Lab Search Results**

**Fig. X**: The tier of confidence of this map for internal use only is **High**. Map represents cetacean sightings from the Whitehead Lab. Absence of a species in the map should be interpreted as an absence of reporting, not as an absence of the species in the area (e.g. field efforts were primarily conducted in the Gully MPA, and adjacent canyons along the edge of the eastern Scotian Shelf, sightings information underrepresents presence of cetaceans as this search does not include acoustic detections). For more information, please visit: https://whiteheadlab.weebly.com/contact.html

#### **WSDB Search Results**
```{r read wsdb and filter SARA & COSEWIC cetacean spp, echo=FALSE, include=FALSE, cache=FALSE}
wsdb_filter <- filter_wsdb(wsdb)
wsdb_filter <- wsdb_filter %>% rename(Scientific_Name = SCIENTIFICNAME)
wsdb_filter <- merge(wsdb_filter, listed_species, by='Scientific_Name')
```

```{r echo=FALSE, results='asis'}
wsdb_plot_locations<-st_as_sf(wsdb_filter, coords = c("LONGITUDE", "LATITUDE"), crs = 4326)
intersect_wsdb <- st_intersection(wsdb_plot_locations,PEZ_poly_st)
x<-as.numeric(nrow(intersect_wsdb))
Common_Name<-intersect_wsdb$Common_Name
Scientific_Name<-intersect_wsdb$Scientific_Name
intersect_wsdb_df<-as.data.frame(cbind(Common_Name, Scientific_Name, row.names = FALSE))

#append columns from listed species 
wsdb_table<-merge(intersect_wsdb_df, listed_species, by='Scientific_Name')
wsdb_table2<-wsdb_table %>% 
  transmute(Common_Name, Scientific_Name, Schedule.status, COSEWIC.status, Wild_Species)
wsdb_table3<- wsdb_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )
Query_output_wsdb<-if(x < 1){
  "There are no records in the WSDB for this search area."
} else {
  "The WSDB contains records of whales in this search area."
}
Query_output_wsdb2<-noquote(Query_output_wsdb)
writeLines(c(Query_output_wsdb2), sep = "\n")
```

**Table 6:** Cetaceans with observations contained in the WSDB within the search polygon area.

```{r echo=FALSE, results='asis'}
WSDB_query_output_table<-if(x < 1){
  ""
} else {
  kable(wsdb_table3[!duplicated(wsdb_table3),], row.names=F) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
WSDB_query_output_table
```

```{r Fig 3 Whale Sightings Database, fig.height=8, fig.width=11}
par(mfrow = c(1,2), mar=c(0, 0, 5.5, 0) + 2)#it goes c(bottom, left, top, right)
wsdb_plot<-site_map_new(PEZ_poly,PEZ_poly_st,site_sf,land_sf,5)+
  geom_point(data = wsdb_filter, aes(x = LONGITUDE, y = LATITUDE, color=Common_Name), 
             size = 3.5)
print(wsdb_plot + scale_colour_manual(values=c("darkgoldenrod1", "chartreuse4", "black", "#00AFBB","#B337A3", "#0827EF", "#EF6408", "#F5A4E7")))

```
**Fig. 3:** The tier of confidence of this map for internal use only is **Low**. Map represents cetacean sightings from the Whale Sightings Database (WSDB) (including sightings of groups of animals) in the search area. Absence of a species in the map should be interpreted as an absence of reporting, not as an absence of the species in the area (e.g. sightings information underrepresents presence of cetaceans as this search does not include acoustic detections). Sightings on land are an indicator that the sighting data have not yet been completely error-checked. 

#### **OBIS Search Results**
```{r, echo=FALSE, include=FALSE, cache=FALSE}
#Select columns of interest from OBIS file
obis_edit<-obis %>% 
  transmute(scientificName, decimalLatitude, decimalLongitude, year,individualCount, rightsHolder, institutionID, institutionCode, collectionCode, datasetID)
 
#Separate CWS/ECCC records
#CWS.ECCC.OBIS_records<-filter(obis_edit, institutionCode=="Canadian Wildlife Service-Atlantic, Environment and Climate Change Canada")
#write.csv(CWS.ECCC.OBIS_records, "CWS_ECCC_OBIS_records.csv")

#Delete WSDB and CWS/ECCC records
obis_edit2<-obis_edit %>%
  filter(! institutionCode =="Canadian Wildlife Service-Atlantic, Environment and Climate Change Canada")
obis_edit3<-obis_edit2 %>%
  filter(! collectionCode =="WHALESITINGS")

obis_plot_locations<-st_as_sf(obis_edit3, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
intersect_obis <- st_intersection(obis_plot_locations,PEZ_poly_st)
x<-as.numeric(nrow(intersect_obis))

#summarize individual counts
#species_sum<-aggregate(intersect_obis$individualCount, by=list(Species=intersect_obis$scientificName), FUN=sum)

#append columns from listed species
intersect_obis<- intersect_obis %>% rename(Scientific_Name=scientificName)

obis_table<-merge(intersect_obis, listed_species, by='Scientific_Name')
obis_table2<-obis_table %>% 
  select(Common_Name, Scientific_Name, Schedule.status, COSEWIC.status, Wild_Species)
obis_table3<- obis_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )
obis_table3$geometry<-NULL
obis_table4<-unique(obis_table3)

Query_output_obis<-if(x < 1){
  "There are no records of priority specues in the OBIS database for the provided search area."
} else {
  "The OBIS database contains records of priority species in the provided search area."
}

Query_output_obis2<-noquote(Query_output_obis)

writeLines(c(Query_output_obis2), sep = "\n")
```

**Table 5:** Priority species with observations contained in the OBIS database within the search polygon area.

```{r echo=FALSE, results='asis'}
OBIS_query_output_table<-if(x < 1){
  ""
} else {
  kable(obis_table4, row.names = F) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
  
}

OBIS_query_output_table

```

```{r Fig 5 OBIS records for cetacean section in PEZ polygon, fig.height=8, fig.width=11}
plot_obis(PEZ_poly,PEZ_poly_st,site_sf,land_sf,intersect_obis,2)

```
**Fig.5:** The tier of confidence of this map is **Medium**. Map represents cetacean sightings from OBIS in the search area. Absence of a species in the map should be interpreted as an absence of reporting, not as an absence of the species in the area (e.g. sightings information underrepresents presence of cetaceans as this search does not include acoustic detections). Sightings on land are an indicator that the sighting data have not yet been completely error-checked.

#### **CWS/ECCC Search Results**
```{r, echo=FALSE, include=FALSE, cache=FALSE}
#Select columns of interest from OBIS file
cws_edit<-cws %>% 
  transmute(scientificName, decimalLatitude, decimalLongitude, year,individualCount, rightsHolder, institutionID, institutionCode, collectionCode, datasetID)
cws_plot_locations<-st_as_sf(cws_edit, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
intersect_cws <- st_intersection(cws_plot_locations,PEZ_poly_st)
x<-as.numeric(nrow(intersect_cws))

#summarize individual counts
#species_sum<-aggregate(intersect_obis$individualCount, by=list(Species=intersect_obis$scientificName), FUN=sum)

#append columns from listed species
intersect_cws<- intersect_cws %>% rename(Scientific_Name=scientificName)

cws_table<-merge(intersect_cws, listed_species, by='Scientific_Name')
cws_table2<-cws_table %>% 
  select(Common_Name, Scientific_Name, Schedule.status, COSEWIC.status, Wild_Species)
cws_table3<- cws_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )
cws_table3$geometry<-NULL
cws_table4<-unique(cws_table3)

Query_output_cws<-if(x < 1){
  "There are no records of priority specues in the CWS/ECCC database for the provided search area."
} else {
  "The CWS/ECCC database contains records of priority species in the provided search area."
}

Query_output_cws2<-noquote(Query_output_cws)

writeLines(c(Query_output_cws2), sep = "\n")
```

**Table 5:** Priority species with observations contained in the CWS_ECCC database within the search polygon area.

```{r echo=FALSE, results='asis'}
CWS_query_output_table<-if(x < 1){
  ""
} else {
  kable(cws_table4, row.names = F) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
  
}

CWS_query_output_table

```

```{r Fig 5 CWS_ECCC records in PEZ polygon, fig.height=8, fig.width=11}
plot_cws(PEZ_poly,PEZ_poly_st,site_sf,land_sf,intersect_cws,2)

```
**Fig.6:** Map showing the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow) with CWS_ECCC records shown as black points.


#### **iNaturalist Search Results**
```{r, echo=FALSE, include=FALSE, cache=FALSE}
#Select columns of interest from OBIS file
inat_edit<-inat %>% 
  transmute(scientific_name, datetime, latitude, longitude, common_name, url)
inat_plot_locations<-st_as_sf(inat_edit, coords = c("longitude", "latitude"), crs = 4326)
intersect_inat <- st_intersection(inat_plot_locations,PEZ_poly_st)
x<-as.numeric(nrow(intersect_inat))

#summarize individual counts
#species_sum<-aggregate(intersect_obis$individualCount, by=list(Species=intersect_obis$scientificName), FUN=sum)

#append columns from listed species
intersect_inat<- intersect_inat %>% rename(Scientific_Name=scientific_name)
inat_table<-merge(intersect_inat, listed_species, by='Scientific_Name')
inat_table2<-inat_table %>% 
  select(Common_Name, Scientific_Name, Schedule.status, COSEWIC.status, Wild_Species)
inat_table3<- inat_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )
inat_table3$geometry<-NULL
inat_table4<-unique(inat_table3)

Query_output_inat<-if(x < 1){
  "There are no records of priority specues in the iNaturalist database for the provided search area."
} else {
  "The iNaturalist database contains records of priority species in the provided search area."
}

Query_output_inat2<-noquote(Query_output_inat)

writeLines(c(Query_output_inat2), sep = "\n")
```

**Table 6:** Priority species with observations contained in the iNaturalist database within the search polygon area.

```{r echo=FALSE, results='asis'}
INAT_query_output_table<-if(x < 1){
  ""
} else {
  kable(inat_table4, row.names = F) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
  
}

INAT_query_output_table

```

```{r Fig 5 INAT records in PEZ polygon, fig.height=8, fig.width=11}
plot_inat(PEZ_poly,PEZ_poly_st,site_sf,land_sf,inat_table,2)

```
**Fig.7:** Map showing the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow) with iNaturalist records shown as black points.

#### **Priority Areas to Enhance Monitoring of Cetaceans Search Results**

```{r echo=FALSE, results='asis'}
#function for overlap
fin_area<-fin_overlap(fin_whale_sf, PEZ_poly_st)
harbour_area<-harbour_overlap(harbour_porpoise_sf, PEZ_poly_st)
humpback_area<-humpback_overlap(humpback_whale_sf, PEZ_poly_st)
sei_area<-sei_overlap(sei_whale_sf, PEZ_poly_st)
cetacean_matrix<-data.frame(Fin_Whale="",Habour_Porpoise="", Humpback_Whale="", Sei_Whale="")
cetacean_matrix[1,1]<-fin_area
cetacean_matrix[1,2]<-harbour_area
cetacean_matrix[1,3]<-humpback_area
cetacean_matrix[1,4]<-sei_area
```

**Table 6:** Results showing if the search area overlap with areas predicted as Priority Areas to Enhance Monitoring of several species of cetaceans (Fin whale, Harbour Porpoise, Humpback Whale and Sei Whale). TRUE = polygon does overlap with priority area; FALSE = polygon does NOT overlap with priority area.  
Important usage considerations: model predictions in this section should not be used on their own for this report, marine spatial planning purposes, and/or to delineate important habitat for cetacean species. Rather, outputs should be used together with other sources of information such as prey distribution, tagging data, detections from acoustic monitoring, other data on cetacean occurrence, and other modeling efforts already available for the area. In fact, the data product summarized below represents an excellent framework in which to properly use the outputs of this section. 


```{r}
kable(cetacean_matrix, align="c") %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

```{r Fig X priority areas for cetaceans, fig.height=8, fig.width=11}
plot_cetaceans_4grid(fin_whale_sf, harbour_porpoise_sf, humpback_whale_sf, sei_whale_sf, PEZ_poly_st, land_sf)
```
**Fig. 8:** The tier of confidence of this map is **Low**. Yellow indicates consolidated Species Distribution Modelling outputs (60-100% relative predicted occurrence rate). These habitats with high suitability are to be interpreted as areas where cetacean monitoring efforts may be prioritized, and results can help direct future survey efforts. Due to the many reasons listed in the discussion section in Gomez et al. 2020, these cetacean modelling outputs do not represent a complete and current distribution of cetaceans in the region. They represent priority areas to direct cetacean monitoring efforts. Their use in marine spatial planning processes should be accompanied by complementary approaches such as the process summarized in the section below. Source: Gomez, C., Konrad, C.M., Vanderlaan, A., Moors-Murphy, H.B., Marotte, E., Lawson, J., Kouwenberg, A-L., Fuentes-Yaco, C., Buren, A. 2020. Identifying priority areas to enhance monitoring of cetaceans in the Northwest Atlantic Ocean. Can. Tech. Rep.Fish. Aquat. Sci. 3370: vi + 103 p. http://waves-vagues.dfo-mpo.gc.ca/Library/40869155.pdf


#### **Search Results for Blue Whale Important Habitat**

Table x. Results showing if the search area overlap with Important areas of Blue Whale feeding, foraging and migration in eastern Canada. TRUE = polygon does overlap with the area; FALSE = polygon does NOT overlap with the area.  

```{r echo=FALSE, results='asis'}
#function for overlap
blue_whale_habitat_overlap(Blue_Whale_shp=Blue_4326b, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for sector
blue_whale_habitat_sector(Blue_Whale_shp=Blue_4326b, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for activity
blue_whale_habitat_activity(Blue_Whale_shp=Blue_4326b, PEZ_poly_st= PEZ_poly_st)
```

```{r echo=FALSE, results='asis'}
#function for months
blue_whale_habitat_months(Blue_Whale_shp=Blue_4326, PEZ_poly_st= PEZ_poly_st)
```

NOTE: Colour code polygons for different activities.  

```{r Fig X PEZ polygon with all maritime blue whale habitat polygons, fig.height=8, fig.width=11}
plot_bw_hab(Blue_4326, PEZ_poly_st, land_sf)

```
**Fig. 9:** The tier of confidence of this map is **High**. Polygons delimit areas in Canadian waters that are important to blue whales for foraging (green) and transit (blue): (1) lower St. Lawrence Estuary – northwestern Gulf of St. Lawrence, (2) Mecatina Trough, (3) south and southwestern Newfoundland, (4) continental shelf edge, (5) Honguedo Strait, (6) Cabot Strait. These six habitat areas identified in the report by Lesage et al. (2018) are considered important for the survival and recovery of blue whales from the western North Atlantic population. Blue whales most likely need to use several of these important habitats to fulfill their biological needs. As a result, access corridors and habitat they connect need to be considered equally important for the population. Source: Lesage, V., Gosselin, J-F., Lawson, J. W., McQuinn, I., Moors-Murphy, H., Plourde, S., Sears. R. et Y. Simard. 2018. Habitats important to blue whales (Balaenoptera musculus) in the Western North Atlantic. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/080: iv + 50 p. <https://waves-vagues.dfo-mpo.gc.ca/Library/40687776.pdf>

```{r Fig X close up of PEZ polygon and bw habitat, fig.height=8, fig.width=11}
plot_bw_hab_zoom(Blue_4326, PEZ_poly_st, land_sf)
```
**Fig. 10:** The tier of confidence of this map is **High**. Polygons delimit areas in Canadian waters that are important to blue whales for foraging (green) and transit (blue): (1) lower St. Lawrence Estuary – northwestern Gulf of St. Lawrence, (2) Mecatina Trough, (3) south and southwestern Newfoundland, (4) continental shelf edge, (5) Honguedo Strait, (6) Cabot Strait. These six habitat areas identified in the report by Lesage et al. (2018) are considered important for the survival and recovery of blue whales from the western North Atlantic population. Blue whales most likely need to use several of these important habitats to fulfill their biological needs. As a result, access corridors and habitat they connect need to be considered equally important for the population. Source: Lesage, V., Gosselin, J-F., Lawson, J. W., McQuinn, I., Moors-Murphy, H., Plourde, S., Sears. R. et Y. Simard. 2018. Habitats important to blue whales (Balaenoptera musculus) in the Western North Atlantic. DFO Can. Sci. Advis. Sec. Res. Doc. 2016/080: iv + 50 p. <https://waves-vagues.dfo-mpo.gc.ca/Library/40687776.pdf>
