# **Search Results for Species at Risk**  

## **National Aquatic Species at Risk Section**  
**Contact:** Carolyn Bakelaar  
**URL:** <http://dfonl7swvgip001.ent.dfo-mpo.ca/Html5Viewer/index.html?viewer=NationalSARMap_EN&LayerTheme=0&locale=en-US#>  

The information below is derived from the *National Aquatic Species at Risk tool* (URL provided above) which uses a Geographic Information System (GIS) to display the distribution, residences and Critical Habitats for aquatic species listed under the Species at Risk Act (SARA). This tool has been developed primarily to facilitate project assessments under the Fisheries Act and the SARA. Here we provide some of the outputs of that tool; however, the optimal source for these open data records is the original tool, which also creates reports.

```{r Set up paths for National Aquatic Species at Risk Section, echo=FALSE, include=FALSE, cache=FALSE}
fl <- list.files(dataPath,".csv")
ClippedCritHab <- st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/ClipCritHab.shp", quiet=TRUE)
#SAR distribution shapefile was clipped from National shapefile as follows:
#sardist<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/Distribution_FGP_fixed_clipped.shp", quiet=TRUE)
#sardist_4326 <- st_transform(sardist, crs = 4326)
#st_write(sardist_4326, "../../../Data/NaturalResources/Species/SpeciesAtRisk/sardist_4326.shp")
sardist<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/sardist_4326.shp", quiet=TRUE)
```
Data products used in this section:  

### **Species At Risk Distribution Range**  

URL: <https://open.canada.ca/data/en/dataset/e0fabad5-9379-4077-87b9-5705f28c490b>  
Tier of confidence: **High**  

The Species at Risk (SAR) Program is responsible for carrying out DFO’s mandate under the Species at Risk Act (SARA) to protect, recover and conserve all listed aquatic SAR in Canada. As part of this mandate, this spatial database has been developed to identify areas in which aquatic species listed under SARA may be found.

Distribution and range information are identified for species listed as Endangered, Threatened or Special Concern under SARA.

Distribution (range) polygons and lines were assembled by regional SARA biologists using the best available information, including COSEWIC status reports, recovery potential assessments, academic literature, and expert opinion. These spatial data support the protection, recovery and conservation of species listed as Endangered, Threatened or Special Concern under SARA. Species distributions are also described and displayed in Recovery Strategies, Action Plans and/or Management Plans.

Discrepancies may exist between the distribution data shown in a species’ SARA recovery document and the current spatial data. Please contact DFO for more information on any data discrepancies.

### **Species at Risk Critical Habitat**  

URL: <https://open.canada.ca/data/en/dataset/db177a8c-5d7d-49eb-8290-31e6a45d786c>  
Tier of confidence: **High**  

The Species at Risk (SAR) Program is responsible for carrying out DFO’s mandate under the Species at Risk Act (SARA) to protect, recover and conserve all listed aquatic SAR in Canada.

Critical habitat is identified for species listed as Endangered or Threatened under the Species at Risk Act (SARA).

Critical habitat is defined under section 2 of SARA as: "the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species' critical habitat in the recovery strategy or in an action plan for the species". Section 49(1)(a) of SARA requires that a species' Recovery Strategy/Action Plan include an identification of the species' critical habitat to the extent possible, based on the best available information, including information provided by the Committee on the Status of Endangered Wildlife in Canada (COSEWIC). SARA makes it illegal to destroy any part of the critical habitat of SAR and may impose restrictions on development and construction.

Critical habitats were assembled by SARA regional biologists and recovery teams. They are designed to support the protection and recovery of species listed as Endangered or Threatened under the Species at Risk Act. They are also described and displayed in species' Recovery Documents and Action Plans.

#### **Species At Risk Distribution Range Search Results**

```{r echo=FALSE, results='asis'}
intersect_dist <- st_intersection(sardist,PEZ_poly_st)
x<-as.numeric(nrow(intersect_dist))
intersect_dist$Common_Nam[intersect_dist$Common_Nam == "Sowerby`s Beaked Whale"] <- "Sowerby's Beaked Whale"
Common_Name<-intersect_dist$Common_Nam
Scientific_Name<-intersect_dist$Scientific
Population<-intersect_dist$Population
Area<-intersect_dist$Waterbody
intersect_dist_df<-as.data.frame(cbind(Common_Name, Scientific_Name, Population, Area))

#append columns from listed species 
dist_table<-merge(intersect_dist_df, listed_species, by='Scientific_Name')
dist_table2<-dist_table %>% 
  transmute(Common_Name, Scientific_Name, Population, Area, Schedule.status, COSEWIC.status, Wild_Species)
dist_table3<- dist_table2 %>% rename(SARA_listing=Schedule.status,
                                     COSEWIC_listing=COSEWIC.status,
                                     Wild_Species_listing=Wild_Species 
                                     )

Query_output_dist<-if(x < 1){
  "The provided polygon does not overlap with species at risk distribution range."
} else {
  "The provided polygon overlaps with species at risk distribution range."
}

Query_output_dist2<-noquote(Query_output_dist)

writeLines(c(Query_output_dist2), sep = "\n")
```

**Table 1:** Species at Risk for which the provided polygon overlaps with their distribution.

```{r echo=FALSE, results='asis'}
Query_output_table<-if(x < 1){
  ""
} else {
  kable(dist_table3[!duplicated(dist_table3),]) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
  
}

Query_output_table

```

#### **Species at Risk Critical Habitat Search Results**

```{r echo=FALSE, results='asis'}
intersect <- st_intersection(ClippedCritHab,PEZ_poly_st)
x<-as.numeric(nrow(intersect))
CommonName<-intersect$Common_Nam
Population<-intersect$Population
Area<-intersect$Waterbody
SARA_status<-intersect$SARA_Statu
intersect_df<-data.table(expand.grid(CommonName = CommonName, Population = Population, Area=Area, SARA_status=SARA_status))
Query_output_crit<-if(x < 1){
  "The provided polygon does not overlap with defined critical habitat."
} else {
  "The provided polygon overlaps with Critical Habitat."
}

Query_output_crit2<-noquote(Query_output_crit)

writeLines(c(Query_output_crit2), sep = "\n")
```
```{r echo=FALSE, results='asis'}
Query_output_table<-if(x < 1){
  ""
} else {
  kable(intersect_df[!duplicated(intersect_df),]) %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

}

Query_output_table2<-noquote(Query_output_table)
Query_output_table2
```

```{r Fig X PEZ polygon with all maritime Critical Habitat polygons, fig.height=8, fig.width=11}
plot_crithab(ClippedCritHab, PEZ_poly_st, land_sf)

```
**Fig. 2:** Map showing Species At Risk Critical Habitat (red), the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow).

```{r Fig X close up of PEZ polygon, fig.height=8, fig.width=11}
plot_crithab_zoom(ClippedCritHab, PEZ_poly_st, land_sf)
```
**Fig. 3:** Map showing Species At Risk Critical Habitat (red), the defined Potential Exposure Zone (PEZ; blue) and proposed project polygon (yellow).
