---
title: "Reproducible Report for Species in Maritimes Region - for Internal DFO use only"
author: Synthesis prepared by the Reproducible Reporting Team, steering committee and advisors.
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: paper
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---
```{r load packages, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(Mar.datawrangling)
#library(Mar.utils)
library(knitr)
library(kableExtra)
library(rgdal)
library(maps)
library(lubridate)
library(raster)
library(RCurl)
library(sf)
library(stringr)
library(ggplot2)
library(data.table)
library(gridExtra)
library(dplyr)
library(stars)
library(ggspatial)
library(tidyverse)
#These next two lines are necessary for the generation of the water mark on all plots.
#install.packages("remotes")
#remotes::install_github("terminological/standard-print-output")
library(standardPrintOutput)
```

```{r source functions, include=FALSE, cache=FALSE}
# the .Rmd file sets the working directory based on where it resides and ignores the wd set by the .RProj file
source("site_map.r")
source("site_map_new.r")
source("filter_wsdb.r")
source("filter_narwc_new.r")
# ## Another option is to load directly from the web and read in using Rcurl
#function_urls <- c("https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/site_map.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/plot_wsdb.r",
#                    "https://raw.githubusercontent.com/AtlanticR/MSP/master/SearchPEZ/code/filter_wsdb.r")
#for (i in function_urls){eval(parse(text = RCurl::getURL(i, ssl.verifypeer = FALSE)),envir=.GlobalEnv)}
source("EBSA.R")
source("leatherback.R")
source("cetacean_priority_areas.R")
source("Blue_Whale_habitat.R")
source("fn_SearchRVData.r")
source("fn_Search_ISDB_Data.r")
source("fn_maps.r")
source("fn_Search_MARFIS_Data.r")
```

```{r Define Site, include=FALSE, cache=FALSE}
#HTML output of code is automatically generated in Catalina's computer: RProjects/MSP/SearchPEZ/code
#output of get.data.R should be manually moved to RProjects/MSP/SearchPEZ/outputs/sitenamexkm  
#AquaSiteName <- "TestGully"
#PEZversion <- "Polygon"
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
UserComments <- "Example - First search effort for suitability of a site for a project planned for 2025. Project is expected to have positive/negative impacts on surrounding area. It is likely that some kind of monitoring will be required during operations. Make sure to follow up on all sections pertaining to a particular taxonomic group as this seems to be a concern of the stakeholders."
#make sure you delete all shp.xml files in the folder to be able to run this code
minYear <- 1970
maxYear <- 2020
```

```{r Set up paths, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
#create a folder in outputs with the name of the site and buffer area search
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
# setwd(dataPath) #not needed?
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
#landGDB <- "../../../Data/Boundaries/Coast50K/Coastline50k.gdb"
#land <- readOGR(dsn=landGDB,layer="Land50k_Maritimes",stringsAsFactors=F)
#land_sf <- st_as_sf(land)
land_sf<-st_read("../../../Data/Boundaries/Coast50K/Coastline50k_SHP/Land_AtlCanada_ESeaboardUS.shp", quiet=TRUE)
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
#PEZ_poly_st<-st_read(../../../Data/Zones/SearchPEZpolygons/PEZ_TestGullyPolygon.shp", quiet=TRUE)
PEZ_poly_st<-st_as_sf(PEZ_poly)
#load data sets that are used in multiple sections
obis<-read.csv("../../../Data/NaturalResources/Species/OBIS_CBIF_iNaturalist/OBIS_MAR_priority_records.csv")
obis<-subset(obis,year>minYear,year<maxYear)
#FFHPP_shp <- st_read("../../../Data/Infrastructure/FFHPP_Referrals/FFHPP_Referrals_0620.shp", quiet=TRUE)
```

```{r Species listed by Wildspecies, COSEWIC and SARA in the Maritime Region, include=FALSE, cache=FALSE}
listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
cetacean_list<-c("Beluga Whale", "Humpback Whale" , "North Atlantic Right Whale", "Fin Whale", "Northern Bottlenose Whale", 
                            "Harbour Porpoise", "Killer Whale", "Blue Whale", "Sei Whale", "Sowerby's Beaked Whale")
other_species_list<-c("Loggerhead Sea Turtle", "Atlantic Walrus", "Harbour Seal Lacs des Loups Marins subspecies", "Leatherback Sea Turtle")
listed_cetacean_species<-subset(listed_species, Common_Name %in% cetacean_list)
listed_other_species<-subset(listed_species, Common_Name %in% other_species_list)
listed_fish_invert_species<-listed_species[ ! listed_species$Common_Name %in% c(other_species_list,cetacean_list), ]
```

## **Habitat Information**

### **Intertidal Vegetation and Rockweed**
Satellite-based Maps of Intertidal Vegetation and Rockweed presence polygons
Contact: <Gordana.Lazin@dfo-mpo.gc.ca>
Last retrieved on: February 17 2021 
Quality Tier: **Medium**
Security level: none.

### **Area-specific Intertidal Vegetation and Rockweed search results**  

"There are no relevant records of predicted intertidal vegetation for this search area" OR "There are There are relevant records of predicted intertidal vegetation for this search area."  

Figure x. Quality Tier: Medium. Security level: none. Map showing the search area defined by the user: location of the proposed project/activity is highlighted in yellow (input polygon), surrounded by an exposure zone or user-defined buffer area in blue. Figure shows overlap between the search are and satellite-based maps of intertidal vegetation and rockweed presence polygons. 