```{r load packages, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
library(easypackages)
libraries("Mar.datawrangling","knitr","kableExtra","rgdal","maps","lubridate","raster","RCurl","sf","stringr","ggplot2","data.table","gridExtra","dplyr",
          "stars","ggspatial","tidyverse","standardPrintOutput")
```
```{r load source functions, include=FALSE, cache=FALSE}
source("site_map.r")
source("site_map_new.r")
source("filter_wsdb.r")
source("filter_narwc_new.r")
source("EBSA.R")
source("leatherback.R")
source("cetacean_priority_areas.R")
source("Blue_Whale_habitat.R")
source("fn_SearchRVData.r")
source("fn_Search_ISDB_Data.r")
source("fn_maps.r")
source("fn_Search_MARFIS_Data.r")
```
```{r Define Site, include=FALSE, cache=FALSE}
AquaSiteName <- "FarmersLedge"
PEZversion <- "4748m"
minYear <- 2002
```
```{r set up paths, include=FALSE, cache=FALSE}
mspPath <- "../"
data.dir = "../../../Data/mar.wrangling"
dataPath <- file.path("../../../Data/outputs",paste0(AquaSiteName,PEZversion))
polyPath <- "../../../Data/Zones/SearchPEZpolygons"
site <- readOGR(polyPath,layer=paste0("Site_",AquaSiteName))
site_sf <- st_as_sf(site)
land_sf<-st_read("../../../Data/Boundaries/Coast50K/Coastline50k_SHP/Land_AtlCanada_ESeaboardUS.shp", quiet=TRUE)
pl <- list.files(polyPath,"*.shp")
pl <- pl[-grep("xml",pl)]
PEZ_poly <- readOGR(file.path(polyPath,pl[grep(paste0("PEZ_",AquaSiteName,PEZversion),pl)]))
PEZ_poly_st<-st_as_sf(PEZ_poly)
```
```{r load SAR species lists, include=FALSE, cache=FALSE}
listed_species<-read.csv("../../../Data/NaturalResources/Species/MAR_listed_species.csv")
cetacean_list<-c("Beluga Whale", "Humpback Whale" , "North Atlantic Right Whale", "Fin Whale", "Northern Bottlenose Whale", "Harbour Porpoise", "Killer Whale", "Blue Whale", "Sei Whale", "Sowerby's Beaked Whale")
other_species_list<-c("Loggerhead Sea Turtle", "Atlantic Walrus", "Harbour Seal Lacs des Loups Marins subspecies", "Leatherback Sea Turtle")
listed_cetacean_species<-subset(listed_species, Common_Name %in% cetacean_list)
listed_other_species<-subset(listed_species, Common_Name %in% other_species_list)
listed_fish_invert_species<-listed_species[ ! listed_species$Common_Name %in% c(other_species_list,cetacean_list), ]
```
```{r Set up paths for National Aquatic Species at Risk Section, include=FALSE, cache=FALSE}
fl <- list.files(dataPath,".csv")
ClippedCritHab <- st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/ClipCritHab.shp", quiet=TRUE)
sardist<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/sardist_4326.shp", quiet=TRUE)
leatherback_shp<-st_read("../../../Data/NaturalResources/Species/SpeciesAtRisk/LeatherBackTurtleCriticalHabitat/LBT_CH_2013.shp", quiet=TRUE)
```

## **Information from National Aquatic Species at Risk Program**  

This section...

### **Species At Risk distribution and Critical Habitat data**  

#### ***DFO Species At Risk distribution (range)***  
Contact: <info@dfo-mpo.gc.ca>  
Link: <https://open.canada.ca/data/en/dataset/e0fabad5-9379-4077-87b9-5705f28c490b>  
Last retrieved on: October 1 2020 from Open Data  
Quality Tier: High  
Security level: none  

#### ***Critical Habitat of Species At Risk***  
Contact: <info@dfo-mpo.gc.ca>  
Link: <https://open.canada.ca/data/en/dataset/db177a8c-5d7d-49eb-8290-31e6a45d786c>  
Last retrieved on: October 1 2020 from Open Data  
Quality Tier: High  
Security level: none  

#### ***Area-specific SAR distribution and critical habitat search results***

*
```{r echo=FALSE, results='asis'}
intersect_dist <- st_intersection(sardist,PEZ_poly_st)
intersect_dist$Common_Nam[intersect_dist$Common_Nam == "Sowerby`s Beaked Whale"] <- "Sowerby's Beaked Whale"
Common_Name<-intersect_dist$Common_Nam
Scientific_Name<-intersect_dist$Scientific
Population<-intersect_dist$Population
Area<-intersect_dist$Waterbody
intersect_dist_df<-as.data.frame(cbind(Common_Name, Scientific_Name, Population, Area))

#append columns from listed species 
dist_table<-merge(intersect_dist_df, listed_species, by='Scientific_Name')
x<-as.numeric(nrow(dist_table))
dist_table2<-dist_table %>% 
  transmute(Common_Name, Scientific_Name, Population, Area, Schedule.status, COSEWIC.status, Wild_Species)
dist_table3<- dist_table2 %>% rename("SARA status"=Schedule.status,
                                     "COSEWIC listing"=COSEWIC.status,
                                     "Wild Species listing"=Wild_Species,
                                     "Common Name"=Common_Name,
                                     "Scientific Name"=Scientific_Name)
Query_output_dist<-if(x < 1){
  "The provided polygon does not overlap with species at risk distribution range."
} else {
  "The provided polygon overlaps with species at risk distribution range."
}
Query_output_dist2<-noquote(Query_output_dist)
writeLines(c(Query_output_dist2), sep = "\n")
```

```{r echo=FALSE, results='asis'}
Query_output_table<-if(x < 1){
  ""
} else {
  kable(dist_table3[!duplicated(dist_table3),], caption= "Quality Tier: High. Security level: none. Species At Risk listed as Endangered, Threatened or Special Concern under the <i>Species At Risk Act</i> for which the search polygon overlaps with their distribution (range). This is <b>not</b> the authoritative source or advice for Species At Risk data. Please access the Species At Risk GIS tool to create an authoritative SARA report.") %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
if(x > 1){Query_output_table}
```

*
```{r echo=FALSE, results='asis'}
intersect <- st_intersection(ClippedCritHab,PEZ_poly_st)
x<-as.numeric(nrow(intersect))
CommonName<-intersect$Common_Nam
Population<-intersect$Population
Area<-intersect$Waterbody
SARA_status<-intersect$SARA_Statu
intersect_df<-data.table(expand.grid(CommonName = CommonName, Population = Population, Area=Area, SARA_status=SARA_status))
Query_output_crit<-if(x < 1){
  "The provided polygon does not overlap with defined critical habitat."
} else {
  "The provided polygon overlaps with Critical Habitat."
}

Query_output_crit2<-noquote(Query_output_crit)
writeLines(c(Query_output_crit2), sep = "\n")
```

```{r echo=FALSE, results='asis'}
Query_output_table<-if(x < 1){
  ""
} else {
  kable(intersect_df[!duplicated(intersect_df),], caption="Quality Tier: High. Security level: none. Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA) for which the search polygon overlaps with their critical habitat. Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the SAR Public Registry (www.sararegistry.gc.ca). This is <b>not</b> the authoritative source or advice for Species At Risk data. Please access the Species At Risk GIS tool to create an authoritative SARA report.") %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
}
Query_output_table2<-noquote(Query_output_table)

if(x > 1){print(Query_output_table2)}
```

```{r Fig X PEZ polygon with all maritime Critical Habitat polygons, fig.height=8, fig.width=11, fig.cap="Quality Tier: High. Security level: none. This map summarizes areas where the search polygon overlaps with critical habitat of Species At Risk listed as Endangered or Threatened under the <i>Species At Risk Act</i> (SARA). Critical habitat is defined under section 2 of SARA as: “the habitat that is necessary for the survival or recovery of a listed wildlife species and that is identified as the species’ critical habitat in the recovery strategy or in an action plan for the species”. Critical Habitat is identified in a species’ recovery strategy or action plan, posted on the SAR Public Registry (www.sararegistry.gc.ca). This is not the authoritative source or advice for Species At Risk data. Please access the Species At Risk GIS tool to create an authoritative report."}
plot_crithab(ClippedCritHab, PEZ_poly_st, land_sf)

```

```{r Fig X close up of PEZ polygon, fig.height=8, fig.width=11, fig.cap="Zoomed view of figure above."}
plot_crithab_zoom(ClippedCritHab, PEZ_poly_st, land_sf)
```