---
title: "Reproducible Report for Species in Maritimes Region - for Internal DFO use only"
author: Synthesis prepared by the Reproducible Reporting Team, steering committee and advisors.
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  bookdown::html_document2:
    number_sections: FALSE
    toc: true
    toc_depth: 5
    theme: journal
bibliography: mybiblio.bib
---
<style>
table {background-color: white !important;color: black !important;}
caption {color: black !important;}
a:link {color: #002080;}
a:visited {color: #008040;}
a:hover {color: #006080;}
</style>

```{r load packages, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="left")
```
```{r, include=FALSE}
# https://bookdown.org/yihui/rmarkdown-cookbook/source-script.html
source("setup_PRG.R", local = knitr::knit_global())
```

# **REPRODUCIBLE REPORTING**    

#### **About this report**    

This document summarizes regional data, including information on species and habitats, available at the time of report creation. …
 
#### **How to use this report?**    

  -  This document

# **SEARCH RESULTS: EMPHASIS ON SPECIES LISTED BY THE <i>SPECIES AT RISK ACT</i>, ASSESSED BY THE COMMITTEE ON THE STATUS OF THE ENDANGERED WILDLIFE IN CANADA (COSEWIC) AND/OR WILD SPECIES**    

```{r echo=FALSE, results='asis'}
List_table<-data.frame("Framework"="", Description = "")
List_table[1,1]<-"SARA"
List_table[2,1]<-"COSEWIC"
List_table[3,1]<-"Wild Species"
```
```{r}
kable(
  List_table,
  align="l", 
  caption = "\\label{tab:List_table}Definitions relevant for species listed by <i>Species At Risk Act</i> (SARA), assessed by the Committee on the Status of Endangered Wildlife in Canada (COSEWIC), or included in the Wild Species listings.",
booktabs = T,
escape = F,
linesep="")%>%
column_spec(1,width = "3cm")%>%
  kable_styling(bootstrap_options = "striped", full_width = T, position = "left")

RVCatch_sf <-  SelectRV_fn(SurveyPrefix, File, studyArea, minYear)
MARFISCatch_sf <- SelectMARFIS_fn(studyArea, minYear)
# Add in a plot
# site_map(PEZPoly_sf,site_sf,land50k_sf,5)
# plot_wsdb(PEZPoly_sf,wsdb_filter,5)

```

#```{r child = 'report_pt2_SAR_dist_crithab_EN_troubleshoot.Rmd'}
#```

#```{r child = 'report_pt3_fish_inverts_EN_troubleshoot.Rmd'}
#```

#```{r child = 'report_pt4_cetaceans_EN_troubleshoot.Rmd'}
#```

#```{r child = 'report_pt5_Habitat_Information_EN_troubleshoot.Rmd'}
#```

#```{r child = 'report_pt6_EBSA_EN_troubleshoot.Rmd'}
#```

#```{r child = 'report_pt7_generic_END_EN.Rmd'}
#```

# **REFERENCES**
[@OpenData_SARrange]
[@WSDB]
[@Whitehead2012]
[@DFO2020NBW_imphabitat]

